### 2016-07-08 paper/icse2017/anatomy.tex
changes in text

\section{Anatomy of \pef{} failure}

\Fix{...}

It is important to note that previous work investigated test flakiness
manifested in sequential executions [cite cite].  For example, a test
case can fail because of a broken test dependency caused when ... or
because of an implicit ordering assumption was broken in a
multi-threaded test (under a sequential execution of the test suite).
Treating this source of flakiness is important but complementary to
this work: parallel execution flakiness can arise even without
sequential execution flakiness.  Discovering and fixing this source of
flakiness is challenging and essential to enable test parallelization.