### 2016-07-08 paper/icse2017/intro.tex

\section{Introduction}

Optimizing regression testing is important.  As software evolves the
size of test sets (both in number and length of tests) increases,
potentially leading to increasing disruption of the development
process caused by late reports of test failures.  This problem
manifests more dramatically at large IT companies (e.g.,\Comment{
  Groupon~\Fix{cite},} Google~\Fix{cite} and Microsoft~\Fix{cite})
where the size of test sets often exceeds the size of the code base.
Several approaches have been proposed to optimize test execution.
Research has focused mostly on test selection~\Fix{cite}, test
reduction~\Fix{cite}, and test prioritization~\Fix{cite}.  Although
these approaches can also be applied in industrial settings, the focus
in industry has been mostly in parallelizing test execution with
server farms, where worker machines from the farm are assigned to
execute distinct subsets of the original test suite.

Given the proliferation of multicore machines it is not surprising
that popular testing frameworks and build systems provide today
support for parallel test execution~\Fix{cite}.  These solutions
attempt to maximize CPU usage in individual machines with the goal of
running tests more efficiently.  In the case of the Java language, it
is possible to explore parallelism across JVMs and within the JVM.  We
conjecture that such lower-level solutions that use commodity hardware
to optimize test execution are also important (compared to
higher-level parallelism from server farms).  For the scenario of
large IT organizations, lower-level parallelization schemes leverage
the computing power on each individual node of a server farm in
addition to the aggregate processing power of the farm.  Lower-level
parallelization fits particularly well smaller organizations
(/projects) with relatively high testing costs but lower budgets.

Unfortunately, these solutions can't be used out-of-the-box:
unrestricted parallel execution of tests can produce non-deterministic
results as developers typically do not provision protection to
concurrent accesses originated from arbitrary program points (\ie{},
tests).  We refer to this problem as Parallel Execution Flakiness
(\pef{}).  To illustrate the importance of parallel test execution and
the problem of \pef{} let's consider the case of the ``core'' module
from the Apache Camel project~\cite{apache-camel-web}.  This module
contains 5,679 test cases, declared in \Fix{NN} test classes.  We ran
those tests in a machine with 16GB of memory and 8 virtual CPUs (4
cores with 2 native threads each).  Sequential test execution takes
24m50s to run this test set.  Execution of the same test set takes
2m28s when we configured parallel execution to spawn a JVM on each CPU
and fork a thread per test class, uniquely allocated to one JVM.  Note
that this is an order of magnitude (10.07x) speedup\Fix{Need to
  understand why this is 10x as opposed to something closer to 7x}.
Unfortunately, due to \pef{}, $\sim$2\% (114 of 5,679) of the tests
fail when executed in parallel.  It is important to notice that the
ratio of failures varies with the project as it depends on factors
such as length of test cases and amount of shared state across tests,
for example.  \pef{} is an important obstacle to enable parallel test
execution: it is important to execute tests efficiently without
sacrificing reliability.  Conceptually, higher parallelization can
result in higher chances of concurrency-related problems.

This paper reports an empirical study to evaluate the impact of
parallelization options on test execution speedup and \pef{}. Based on
these results, we proposed different low-cost synchronization
strategies to avoid \pef{}.

%%  LocalWords:  parallelization multicore JUnit TestNG NUnit XXm YYm
%%  LocalWords:  Groupon parallelizing multi