### 2016-07-08 paper/icse2017/prelim-study.tex

\section{Preliminary Study}

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    \multirow{2}{*}{\emph{Subject}} & \multirow{2}{*}{\emph{\# Tests}} &  \multicolumn{2}{c|}{\emph{Time}}\\
    \cline{3-4}
    & & Seq & Par \\
    \hline
    jgit & 3364 & 04:12 min & 53.003 s \\
    \hline
    jetty (client) & 5679 & 04:15 min & 18.752 s \\
    \hline
    camel (core) & 527 & 24:50 min & 02:28 min \\
    \hline
  \end{tabular}
  \caption{\label{table:cost}...}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    \emph{Subject} & Fail-Par. & Fail-Ind. \\
    \hline
    jgit & 239 & 4 \\
    \hline
    jetty (client) & 232 & - \\
    \hline
    camel (core) & 142 & - \\
    \hline
  \end{tabular}
  \caption{\label{table:failures}...}
\end{table}