### 2016-07-08 paper/icse2017/anatomy.tex
polish text

\section{Anatomy of a \pef{} failure}

\Fix{...}

It is important to note that previous work investigated test flakiness
manifested in sequential executions [cite cite].  For example, a test
case can fail because an implicit ordering assumption from a
multi-threaded test was broken under a sequential execution of the
test suite.  Treating this source of flakiness is important but
orthogonal to this work: parallel execution flakiness can arise even
without sequential execution flakiness.