### 2016-10-11 paper/conclusions.tex
Renamed paper directory and updated .gitignore file

\section{Conclusions}

\Fix{...}

These tools can be used out-of-the-box to support \emph{system}
testing of applications that have been designed to handle concurrent
user accesses\cite{selenium-hq}.