### 2017-01-06 paper/codes/discovery.tex
adjusted codes

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    showstringspaces=false
}
\begin{lstlisting}[language=Bash]
for subj in $SUBJECTS; do
    subject_path="$SUBJECTS_HOME/$subj"
    build_files=`find $subject_path -name pom.xml`
    for build_file in $build_files; do
        output=`cat $build_file | grep "parallel\|forkMode\|forkCount"`
        if [ -z $output ]; then echo $build_file; fi
done
\end{lstlisting}
    \caption{\label{fig:discovery-step} Bash script to filter
    build files (\ie, \emph{pom.xml}) with potentially parallelization
    settings. We first find all build files from a subject (line 3).
    Then, we iterate over each path (line 4), list the file content
    (\CodeIn{cat} command), and redirect the output to filter only the
    lines that contain any of the keywords \CodeIn{parallel},
    \CodeIn{forkMode}, or \CodeIn{forkCount} (\CodeIn{grep} command).
    If the output is not empty (line 7), we print the path from the
    current \emph{pom.xml} file to inspect it later.}
\end{figure}