### 2017-01-06 paper/codes/evaluation.tex

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    showstringspaces=false
}
\begin{lstlisting}[language=Bash]
MAVEN_IGNORES="-Dmaven.javadoc.skip=true -Drat.skip=true \
    -Djacoco.skip=true -Dcheckstyle.skip=true \
    -Dfindbugs=true -Dcobertura.skip=true"

for subj in $SUBJECTS; do
    cd $SUBJECTS_HOME/$subj
    mvn clean install -DskipTests $MAVEN_IGNORES
    mvn dependency:go-offline
    mvn -o test $MAVEN_IGNORES &> $LOG
    cat $LOG | grep "\[INFO\] Total time:"
done
\end{lstlisting}
\caption{\label{fig:mvn-execution} Bash script to evaluate the time cost of
    test suites. For each subject, we compile the source and test files, fetch
    all dependencies, and execute the tests in offline mode ignoring
    non-related tasks. Unnecessary commands are omitted for simplicity.}
\end{figure}