### 2017-01-08 paper/codes/discovery.tex
Pass on text

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    showstringspaces=false
}
\begin{lstlisting}[language=Bash]
for subj in $SUBJECTS; do
    subject_path="$SUBJECTS_HOME/$subj"
    build_files=`find $subject_path -name pom.xml`
    for build_file in $build_files; do
        output=`cat $build_file | grep "parallel\|forkMode\|forkCount"`
        if [ -z $output ]; then echo $build_file; fi
done
\end{lstlisting}
    \caption{\label{fig:discovery-step} Bash script to filter build
    files for inspeciton: line 3 lists all \emph{pom.xml} files from a
    subject; for each path, the  \CodeIn{cat} command lists the file
    content and uses \CodeIn{grep} to filter only the matching lines
    with the keywords \CodeIn{parallel}, \CodeIn{forkMode}, or
    \CodeIn{forkCount} (lines 5-6); if the output is not empty (line
    7), we print the current path.}
\end{figure}