### 2017-04-03 paper/modes.tex
adjusting examples

\newcommand{\Seq}{C0}
\newcommand{\SeqClassParMeth}{C1}
\newcommand{\ParClassSeqMeth}{C2}
\newcommand{\ParClassParMeth}{C3}
\newcommand{\Fork}{F}
\newcommand{\ForkSeq}{\Fork{}\Seq{}}
\newcommand{\ForkParMeth}{\Fork{}\SeqClassParMeth{}}

\section{Parallel Execution of Test Suites}
\label{sec:modes}

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.35\textwidth]{figs/parallel-levels.pdf}
  \vspace{-1ex}
  \caption{\label{fig:levels}Levels of parallelism.}
\end{figure}

Figure~\ref{fig:levels} illustrates the different levels where
parallelism in test execution can be obtained.
The highest level indicates
parallelism obtained through different machines on the
network.  For instance, using virtual machines from a cloud service to
distribute test execution.  The lowest levels denote parallelism
obtained within a single machine.  These levels are complementary:
the lowest levels leverage the computing power of server
nodes whereas the highest level leverages the aggregate processing
power of the farm.

This paper focuses on low-level parallelism, where computation can be
offloaded at different CPUs within a machine and at different threads
within each CPU.  This form of parallelism is enabled through build
systems (spawning processes in different CPUs) and testing frameworks
(spawning threads in one given CPU).  It is important to note that a variety
of testing frameworks provide today support for parallel test
execution (e.g., JUnit~\cite{junit-org}, TestNG~\cite{testng}, and
NUnit~\cite{nunit}) as to benefit from the available power of popular multi-core processors.
In the following, we elaborate relevant features of testing frameworks
and build systems.  We focused on Java, Maven, and JUnit but the
discussion can be generalized to other language and tools.

\subsection{Testing Frameworks}
\label{sec:frameworks}

The list below shows the choices to control parallelism within one
Java Virtual Machine (JVM), which runs in a single CPU process.  These
options are offered by the testing framework, \eg{}, JUnit.

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~No parallelism is involved.
\item
    \textbf{Sequential classes; parallel methods
      (\SeqClassParMeth).}~This configuration corresponds to running
    test classes sequentially, but running test methods from those
    classes concurrently.
\item
    \textbf{Parallel classes; sequential methods
      (\ParClassSeqMeth{}).}~This configuration corresponds to running
    test classes concurrently, but running test methods sequentially.
\item
    \textbf{Parallel classes; Parallel methods
      (\ParClassParMeth).}~This configuration corresponds to the
    combination of \SeqClassParMeth{} and \ParClassSeqMeth{}.  
\end{itemize}

%% The configuration \Seq{} is a proper fit for short-running test suites
%% since this is the default configuration and the execution provides
%% feedback in an acceptable time. However, for costly suites, it is
%% impractical for developers to wait long-running executions to take
%% action.  In this scenario, the remaining configuration may help to
%% amortize the execution cost depending on how the test suite was
%% designed. For instance,

Notice that an important aspect in deciding which configuration to use
(or in designing new test suites) is the possibility of race
conditions on shared data during execution. For instance,
configuration \ParClassSeqMeth{} may be preferable over
\SeqClassParMeth{} when it is clear that several test methods in one
class perform operations involving one given variable that is visible
statically or within the scope of the test class.

Considering efficiency, the configuration \SeqClassParMeth{} may
speedup the execution of test classes composed of small test methods
involving variables that is visible within the scope of the method.
Similarly, the configuration \ParClassSeqMeth{} may speedup
test suites composed by several test classes with few test methods.
Finally, the configuration \ParClassParMeth{} does not impose any 
sequential execution but it is more likely to manifest data-races as 
methods from different classes run in parallel.

\subsection{Build Systems}
\label{sec:builder}

The list below shows the choices to control parallelism through the
build system, \eg{}, Maven.  The build system can spaw multiple JVMs,
each running on its own OS process on a given CPU and handling a
partition of the test set.

\begin{itemize}
\item
    \textbf{Forked classes; sequential methods (\ForkSeq).}~This
    configuration spawns multiple JVMs, assigning a partition of the
    test set to each JVM.  Test methods run sequentially within each
    JVM.
\item
    \textbf{Forked classes; parallel methods (\ForkParMeth).}~In
    addition to forking multiple JVMs, as \ForkSeq{} does, this
    configuration runs tests concurrently, as \SeqClassParMeth{} does.
\end{itemize}

Several build systems support the execution of test classes in forked
processes running a JVM. In the configuration \ForkSeq{}, each spawned
process runs a different test class at time and the builder merges the
results from each execution. In the configuration \ForkParMeth{}, the
build system forwards settings to the underlying testing framework to
enable parallel execution within each process in addition to running
test classes on different processes.  However, some configurations
(\eg, \ParClassSeqMeth) may not take effect since only one test class
can run in a forked process at time.  To the best of our knowledge, we
are unaware of any build system capable of running multiple classes at
time within a forked process.

%%  to
%% define tasks related to the project building and these tasks are
%% performed several plugins.This file
%% contains defintions of tasks, which are implemented by a collection of
%% plugins.

%% \Jbc{falar do surefire e explicar a
%% configuracao ilustrada - Surefire levanta uma JVM por core e cria uma
%% pool de 5 threads em cada JVM para executar testes em paralelo.}

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    morekeywords={parallel, threadCount, perCoreThreadCount,
    forkCount, reuseFork},
    deletekeywords={true}
}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <forkCount>1C</forkCount>
        <parallel>methods</parallel>
        <threadCount>5</threadCount>
    </configuration>
</plugin>
\end{lstlisting}
    \caption{\label{fig:surefire} Configuration \ForkParMeth{} on Maven.}
\end{figure}

\subsubsection*{Example}~Figure~\ref{fig:surefire} shows a
fragment of a Maven configuration file, known as \pomf{}, highlighting
options to run tests using the parallel execution mode \ForkParMeth{}.
With this configuration, Maven~---~through its test plugin
Surefire~\cite{maven-surefire-plugin}~---~forks one JVM per core
(\CodeIn{forkCount} parameter) and uses five threads
(\CodeIn{threadCount} parameter) to run methods (\CodeIn{parallel}
parameter) within each forked JVM.

%%  LocalWords:  parallelization CPUs JUnit TestNG NUnit multi JVM de
%%  LocalWords:  JVMs falar explicar configuracao ilustrada levanta
%%  LocalWords:  uma por cria cada executar paralelo escapeinside