### 2017-04-03 paper/codes/evaluation.tex

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    showstringspaces=false
}
\begin{lstlisting}[language=Bash]
MAVEN_IGNORES="-Dmaven.javadoc.skip=true -Drat.skip=true \
    -Djacoco.skip=true -Dcheckstyle.skip=true \
    -Dfindbugs=true -Dcobertura.skip=true"

for subj in $SUBJECTS; do
    cd $SUBJECTS_HOME/$subj
    mvn clean install -DskipTests $MAVEN_IGNORES
    mvn dependency:go-offline
    mvn -o test $MAVEN_IGNORES &> $LOG
    cat $LOG | grep "\[INFO\] Total time:"
done
\end{lstlisting}
\caption{\label{fig:mvn-execution} Bash script to measure time cost of
  test suites. For each subject, we compile the source and test files,
  fetch all dependencies, and execute the tests in offline mode
  ignoring non-related tasks. Test-unrelated tasks are omitted.\Jbc{these results will be updated}}
\end{figure}