### 2017-05-10 paper/lessons.tex
another pass

\section{Lessons Learned}

\vspace{1ex}\noindent{}\textbf{The Good.}\Fix{...}

\vspace{1ex}\noindent{}\textbf{The Bad.}\Fix{...}

\vspace{1ex}\noindent{}\textbf{The Ugly.}\Fix{...}