### 2017-05-10 paper/threats.tex

\section{Threats to Validity}

\Fix{To appear... Elaborate CONSTRUCT, INTERNAL and EXTERNAL threats.
I'm going to be writing this ASAP as I finish to present the results
from RQ6}