### 2017-05-11 paper/threats.tex
minor tweaks

\section{Threats to Validity}

\newcommand{\QA}{QA}
\newcommand{\dockerfile}{dockerfile}
\newcommand{\SO}{StackOverflow}

The main threats to validity of this study are the following.
\textit{External Validity:} Generalization of our findings is limited
to our selection of subjects, testing framework, and build system.  To
mitigate that issues, we selected subjects according to an objective
criteria, described in Section~\ref{sec:subjects}.  It remains to
evaluate the extent to which our observations would change when using
different testing frameworks and build systems.  \Mar{Copied from
  Fabrizio+Luis paper.  Use as reference...$\rightarrow$}
\textit{Internal Validity:} Our results could be influenced by
unintentional mistakes made by humans involved in this study.  For
example, students were involved in a user study whereas developers
manually categorized questions in difficulty levels and elaborated
\dockerfile{}s.  All those tasks could introduce bias.  We used Card
Sorting~\cite{lorr1983cluster} to mitigate the problem of incorrectly
categorizing questions.  To make sure the scripts were correct,
developers were instructed to strictly follow the instruction from
\QA{} post preferred answers to reproduce corresponding problems.  We
also encouraged developers to do their best to reproduce as many
questions as possible.  As for the answer of students in the user
study, we analyzed their answers carefully, comparing them with the
solution prepared by the instructors.  It is important to note that
all artifacts produced during this study are publicly available for
scrutiny.  \textit{Construct Validity:} We considered a number of
metrics in this study that could influence some of our
interpretations.  For example, we used metrics of document similarity
to assess how (dis)similar the \dockerfile{}s produced by developers
are.  To mitigate the bias associated with metric selection we used
multiple metrics and confirmed that the similarity was very high as to
not compromise corresponding conclusions.