### 2017-05-13 paper/threats.tex
threats

\section{Threats to Validity}

The main threats to validity of this study are the following.
\textit{External Validity:} Generalization of our findings is limited
to our selection of subjects, testing framework, and build system.  To
mitigate that issue, we selected subjects according to an objective
criteria, described in Section~\ref{sec:subjects}.  It remains to
evaluate the extent to which our observations would change when using
different testing frameworks and build systems.  \textit{Internal
  Validity:} Our results could be influenced by unintentional mistakes
made by humans who interpreted survey data and implemented scripts to
collect and analyze data. All those tasks could introduce bias.  To
mitigate those threats, we worked in pairs as to increase chances of
capturing unintentional mistakes.  \textit{Construct Validity:} We
considered a number of metrics in this study that could influence some
of our interpretations.  For example, we measured number of test cases
per suite, distribution of test costs in a suite, time to run a suite,
etc.  In principle, these metrics may not reflect the main problems
associated with test efficiency.\Mar{...elaborate...}

%%  LocalWords:  QA dockerfile StackOverflow