### 2017-05-13 paper/threats.tex
pass on rq7 (not done); going back to rq6

\section{Threats to Validity}

The main threats to validity of this study are the following.
\textit{External Validity:} Generalization of our findings is limited
to our selection of subjects, testing framework, and build system.  To
mitigate that issue, we selected subjects according to an objective
criteria, described in Section~\ref{sec:subjects}.  It remains to
evaluate the extent to which our observations would change when using
different testing frameworks and build systems.  \textit{Internal
  Validity:} Our results could be influenced by unintentional mistakes
made by humans who interpreted survey data and implemented scripts and
code to collect and analyze data. For example, we developed JUnit
runners to reproduce Maven's parallel configurations.  All those tasks
could introduce bias.  To mitigate those threats, we worked in pairs
as to increase chances of capturing unintentional mistakes.
\textit{Construct Validity:} We considered a number of metrics in this
study that could influence some of our interpretations.  For example,
we measured number of test cases per suite, distribution of test costs
in a suite, time to run a suite, etc.  In principle, these metrics may
not reflect the main problems associated with test
efficiency.\Mar{...elaborate...}

%%  LocalWords:  QA dockerfile StackOverflow