### 2017-08-15 paper/codes/evaluation.tex
Uploading preliminar updates

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    showstringspaces=false
}
\begin{lstlisting}[language=Bash]
MAVEN_SKIPS="-Drat.skip=true -Dmaven.javadoc.skip=true \
    -Djacoco.skip=true -Dcheckstyle.skip=true \
    -Dfindbugs.skip=true -Dcobertura.skip=true \
    -Dpmd.skip=true -Dcpd.skip=true"

for subj in $SUBJECTS; do
    cd $SUBJECTS_HOME/$subj
    mvn clean dependency:go-offline
    mvn test-compile install -DskipTests $MAVEN_SKIPS \
        &> compile.log
    mvn test -o -fae $MAVEN_SKIP &> testing.log
    cat testing.log \
        | grep --text "\[INFO\] Total time:" \
        | tail -n 1
done
\end{lstlisting}
\caption{\label{fig:mvn-execution} Bash script to measure time cost of
  test suites. For each subject, we fetch all dependencies,
  compile the source and test files, and execute the tests in
  offline mode ignoring non-related tasks. Test-unrelated tasks
  are omitted.}
\end{figure}