### 2017-08-15 paper/threats.tex

\section{Threats to Validity}

The main threats to validity of this study are the following.
\textit{External Validity:} Generalization of our findings is limited
to our selection of subjects, testing framework, and build system.  To
mitigate that issue, we selected subjects according to an objective
criteria, described in Section~\ref{sec:subjects}.  It remains to
evaluate the extent to which our observations would change when using
different testing frameworks and build systems.  \textit{Internal
  Validity:} Our results could be influenced by unintentional mistakes
made by humans who interpreted survey data and implemented scripts and
code to collect and analyze data. For example, we developed JUnit
runners to reproduce Maven's parallel configurations and implemented
several scripts to automate our experiments.  All those tasks
could introduce bias in our results.  To mitigate those threats, we
worked in pairs as to increase chances of capturing unintentional
mistakes. Also, we re-executed our experiments on different machines
with different hardware to ensure that our results are consistent.
\textit{Construct Validity:} We considered a number of metrics in this
study that could influence some of our interpretations.  For example,
we measured number of test cases per suite, distribution of test costs
in a suite, time to run a suite, etc.  In principle, these metrics may
not reflect the main problems associated with test
efficiency.
 \Jbc{remember to mention failing tests and how we mitigate
this threat}

%%  LocalWords:  QA dockerfile StackOverflow