### 2017-08-30 paper/threats.tex
Moved content about flaky tests from intro to sec III

\section{Threats to Validity}

The main threats to validity of this study are the following.
\textit{External Validity:} Generalization of our findings is limited
to our selection of subjects, testing framework, and build system.  To
mitigate that issue, we selected subjects according to an objective
criteria, described in Section~\ref{sec:subjects}.  It remains to
evaluate the extent to which our observations would change when using
different testing frameworks and build systems.
\textit{Internal Validity:} Our results could be influenced by
unintentional mistakes made by humans who interpreted survey data and
implemented scripts and code to collect and analyze data.
For example, we developed JUnit runners to reproduce Maven's parallel
configurations and implemented several scripts to automate our
experiments (\eg, run tests and detect parallelism enabled by default
in the subjects).
All those tasks could introduce bias in our results.
To mitigate those threats, we worked in pairs as to increase chances
of capturing unintentional mistakes.
\textit{Construct Validity:} We considered a number of metrics in this
study that could influence some of our interpretations.  For example,
we measured number of test cases per suite, distribution of test costs
in a suite, time to run a suite, etc.  In principle, these metrics may
not reflect the main problems associated with test
efficiency.

%%  LocalWords:  QA dockerfile StackOverflow