### 2017-10-13 paper/evaluation.tex
Fixed overwritten latex rules

\section{Setup and Replication}
\label{sec:setup}

To run our experiments, we used a Core i7-4790 (3.60 GHz) Intel processor
machine with eight virtual CPUs (four cores with two native threads each) and
16GB of memory, running Ubuntu 14.04 LTS Trusty Tahr (64-bit version).  We
used\Comment{ Git,} Java 8 and Maven 3.3.9 to build projects and run test
suites. To process test results and generate plots we used Python\Comment{ 3.4},
Bash, R and Ruby\Comment{ 2.3}.  All source artifacts are publicly available for
replication on our website~\cite{ourwebpage}.  This includes supporting
scripts\Comment{ (\eg, the scripts to run the tests and generate raw analysis
data)} and the full list of projects.

\section{Evaluation}
\label{sec:eval}

We pose the following research questions, organized by the dimensions
of analysis we presented in Section~\ref{sec:intro}.

%% Feasibility
\newcommand{\numRQFeasibilityOne}{RQ1}
\newcommand{\RQFeasibilityOne}{How prevalent are time-consuming
  test suites\Comment{ in open-source projects}?}

\newcommand{\numRQFeasibilityTwo}{RQ2}
\newcommand{\RQFeasibilityTwo}{How is time distributed across test cases?}

%% Adoption
\newcommand{\numRQAdoptionOne}{RQ3}
\newcommand{\RQAdoptionOne}{How popular is test suite
  parallelization\Comment{ in open-source projects}?}

\newcommand{\numRQAdoptionTwo}{RQ4}
\newcommand{\RQAdoptionTwo}{What are the main reasons that prevent developers
  from using test suite parallelization?}

%% Speedups
\newcommand{\numRQSpeedupOne}{RQ5}
\newcommand{\RQSpeedupOne}{What are the speedups obtained with parallelization
  (in projects that actually use it)?}

\newcommand{\numRQSpeedupTwo}{RQ6}
\newcommand{\RQSpeedupTwo}{How test execution scales with the number of
  available CPUs?}

%% Tradeoffs
\newcommand{\numRQIssuesOne}{RQ7}
\newcommand{\RQIssuesOne}{How parallel execution configurations affect testing
  costs and flakiness?}

\setlist[itemize]{leftmargin=1em}
\begin{itemize}
\item Feasibility
  \begin{itemize}
  \item \textbf{\numRQFeasibilityOne.} \RQFeasibilityOne
  \item \textbf{\numRQFeasibilityTwo.} \RQFeasibilityTwo    
  \end{itemize}
\item Adoption
  \begin{itemize}
  \item \textbf{\numRQAdoptionOne.} \RQAdoptionOne    
  \item \textbf{\numRQAdoptionTwo.} \RQAdoptionTwo
  \end{itemize}
\item Speedups
  \begin{itemize}
  \item \textbf{\numRQSpeedupOne.} \RQSpeedupOne
  \item \textbf{\numRQSpeedupTwo.} \RQSpeedupTwo
  \end{itemize}      
\item Tradeoffs
  \begin{itemize}
  \item \textbf{\numRQIssuesOne.} \RQIssuesOne    
  \end{itemize}
\end{itemize}

\subsection{Feasibility}
\label{sec:rqA}
\label{sec:rqB}

\begin{itemize}
  \item \numRQFeasibilityOne{}. \textbf{\RQFeasibilityOne}
\end{itemize}

To evaluate prevalence of projects with time-consuming test suites, we
considered the \numSubjs{} projects, appearing in 
Figure~\ref{fig:subjects}.  Figure~\ref{fig:mvn-execution} illustrates
the script we used to measure time.

We took the following actions to isolate our environment from
measurement noise.
First, we observed that some test tasks called test-unrelated tasks
(\eg, \emph{javadoc} generation and static analyses) that could
interfere in our time measurements.
To address that potential issue, we inspected Maven execution logs
from a sample including a hundred projects prior to running the script
from Figure~\ref{fig:mvn-execution}.
The tasks we found were ignored from execution (lines 1-4).
Furthermore, we configured
our workstation to only run essential services as to avoid noise from unrelated OS events.
The machine was dedicated to our experiments and we
accessed it via SSH. In addition, we configured the \CodeIn{isolcpus}
option from the Linux Kernel \cite{linux-kernel} to isolate six
virtual CPUs to run our experiments, leaving the remaining CPUs to run
OS processes~\cite{isolcpus-use}.  The rationale for this decision is
to prevent context-switching between user processes (running the
experiment) and OS-related processes.  Finally, to make sure our
measurements were fair, we compared timings corresponding to the
sequential execution of tests using Maven with that obtained with
JUnit's default \CodeIn{JUnitCore} runner, invoked from the command
line.  Results were very close.
The main loop (lines 6-15) of the script in
Figure~\ref{fig:mvn-execution} iterates over the list of subjects and
invokes Maven multiple times\Comment{ to isolate cost of running
tests} (lines 8-11).  It first makes all dependencies available locally
(line 8), compiles the source and test files (line 9), and then runs
the tests in offline mode as to skip the package update task, enabled
by default (line 11). After that, we used a regular expression on
the output log to find elapsed times (line 12-14).

\input{codes/evaluation}

\begin{figure}[ht]%
  \centering
  \subfigure[\label{fig:rq1-barplot}]{
  \includegraphics[height=4cm,width=3cm]{results/barplot-timecost.pdf}
  }
  \qquad
  \subfigure[\label{fig:rq1-boxplot}]{
    \includegraphics[height=4cm,width=4.2cm]{results/boxplot-timecost.pdf}
  }%
  \caption{(a)\Comment{ Testing time grouped by time cost ($t$): short run
    ($t<1m$), medium run ($1m \le t < 5m$), and long run ($5m \le{}
    t$);}~Number of projects in each cost group and
    (b)~Distribution of running times per cost group.}
\vspace{-5mm}
\end{figure}

\Comment{We followed a similar methodology to group projects by time as
Gligoric~\etal{}~\cite{gligoric-etal-issta2015} in their work on
regression test selection.}
\Comment{ and added the \medg{} group due to the variability of the
time cost from subjects out of the \shortg{} group}

We ran the test suite for each subject three times, reporting averaged
execution times in three ranges: tests that run within a minute
(\shortg{}), tests that run in one to five minutes (\medg{}), and
tests that run in five or more minutes (\longg{}). Figure~\ref{fig:rq1-barplot} shows the number of projects in
each group.  As expected, \longg{} and \medg{} projects do not occur
as frequently as \shortg{} projects.  However, they do occur in
relatively high numbers.
Figure~\ref{fig:rq1-boxplot} shows the distribution of execution time of
test suites in each of these groups.
Note that the y-ranges are different.
The distribution associated with the \shortg{} group is the most
unbalanced (right skewed)\Comment{ with outliers closed to the \medg{}
group}.
The test suites in this group ran in 15 or less seconds for
over 75\% of the cases.\Comment{  Such scenarios constitute the majority of the
cases we analyzed.} Considering the groups \medg{} and \longg{},
however, we found many costly executions.  Nearly 75\% of the projects
from the \medg{} group take 3.5 or more minutes to run and nearly 75\% of
the projects from the \longg{} group take $\sim$20 minutes to run.  We
found cases in the \longg{} group were execution takes more than 50 minutes
to complete, as can be observed from the outliers in the boxplot.

It is important to note that we under-estimated running times as we missed test modules not enabled for
execution in the root \emph{pom.xml}.\Comment{Some projects may omit long-running tests on their default
execution.} For instance, the project \CodeIn{apache.maven-surefire}
runs all unit tests in a few seconds.  According to our criteria, this
project is classified as \shortg{} but a closer look reveals
that only smoke tests are executed in this project by default.
In this project, integration and system tests, which take longer to run, are only accessible via
custom parameters, which we do not handle in our experimental setup.
We enabled such parameters for this specific project and observed that
testing time goes to nearly 30 minutes.  For simplicity, we considered
only the tests executed by default.
From the \numSubjs{} testable projects, \numSubjsPass{} successfully
executed all tests and \numSubjsFail{} reported some test failure.
From these \numSubjsFail{} subjects, only 11 subjects
have more than 5\% of failing tests (7.3\% on average).
\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQFeasibilityOne{}:}~\emph{We conclude that
    time-consuming test suites are relatively frequent in
    open-source projects.  We found that \percentMedLongRunning{} of
    the \numSubjs{} projects we analyzed (\ie{}, nearly 1 in every 4
    projects) take at least 1 minute to run and
    \percentLongRunning{} of them take at least 5 minutes to run.\Comment{
      (\ie, \numMedLong{} projects from \medg{} and \longg{}).}}
\end{minipage}
}
\end{center}

\begin{itemize}
  \item \numRQFeasibilityTwo. \textbf{\RQFeasibilityTwo}
\end{itemize}

\begin{figure}[t!]
  \centering
  \includegraphics[width=.48\textwidth]{results/testcost-distribution.pdf}
  \caption{\label{fig:time-distributions}Distribution of test case time per project.}%
  \vspace{-5mm}
\end{figure}

Section~\ref{sec:rqA} showed that medium and long-running projects are
not uncommon, accounting to nearly \percentMedLongRunning{} of the
\numSubjs{} projects we analyzed.  Research question \numRQFeasibilityTwo{}
measures the distribution of test costs in test suites.\Comment{ as to estimate
potential of obtaining speedups with parallelization.}  In
the limit, if cost is dominated by a single test from a large test
suite, it is unlikely that parallelization will be beneficial as a
test method is the smallest working unit in test frameworks.
Figure~\ref{fig:time-distributions} shows the time distribution of
individual test cases per project.
We observed that the average median times (see dashed horizontal red
lines) were small, namely 0.08s
for \medg{} projects and 0.16s for \longg{} projects, and the standard deviations associated with each distribution were
relatively low.\Comment{ Figure~\ref{fig:sd} shows the number of
  projects within specific ranges of $\sigma$ values.} High values of
$\sigma$ are indicative of CPU monopolization. We found only a small number
of those. The highest value of $\sigma$ occurred in
\CodeIn{uber\_chaperone}, a project from the \longg{} group.
This project contains only 26 tests, 17 of which take less than 0.5s
to run, one of which takes nearly 3s to run, two of which take nearly
11s to run, four of which takes on average 3m to run, and two of which
take $\sim$8m to run.
For this project, 98.4\% of the execution cost is dominated by 20\% of
the tests; without these two costly tests this project would have been
classified as short-running.
We did not find other projects with such extreme time monopolization
profile.
Project \CodeIn{facebookarchive\_linkbench} is also classified as
long-running and has the second highest value of $\sigma$.
For this project, however, cost is distributed more smoothly across 98
tests, of which 8 (8.1\%) take more than 1s to run with the rest of
the tests running faster.

\begin{figure}[t!]%
  \centering
  \subfigure[\label{fig:size-testsuites}]{
  \includegraphics[height=4.25cm, width=2.1cm]{results/boxplots-testcases.pdf}
  }
  \qquad
  \subfigure[\label{fig:scattercost}]{
    \includegraphics[height=4.25cm]{results/scatter-testcost.pdf}
  }%
  \caption{\label{fig:time-versus-size}(a) Size of test suites; (b)
    Size versus running time of test suites.}%
  \vspace{-5mm}
\end{figure}

Figure~\ref{fig:size-testsuites} shows the difference in the
distribution of test suite sizes across groups.  This figure indicates
that long projects have a higher median and much higher average number of test cases.
Furthermore, we noted a strong positive correlation between running
time and number of test on projects in the \longg{} group.
Considering the \medg{} group, the correlation between these two
variables was weak.
Figure~\ref{fig:scattercost} illustrates the regression lines between
these the variables test suite cost and number of test cases.
To sum, we observed that for projects with long-running test suites
running time is typically
justified by the number of test cases as opposed to the cost of individual test cases.

\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQFeasibilityTwo{}:}~\emph{Overall, results indicate that
  projects with a very small number of tests monopolizing end-to-end
  execution time were rare. Time most often is distributed evenly
  across test cases.}
\end{minipage}
}
\end{center}

\subsection{Adoption}
\label{sec:rqC}
\label{sec:rqE}

\begin{itemize}
    \item \numRQAdoptionOne. \textbf{\RQAdoptionOne{}}
\end{itemize}

To answer \numRQAdoptionOne{} we used projects from the \medg{} and
\longg{} groups where parallelization can be more helpful. We used a
dynamic and a static approach to find manifestations of
parallelism. We discuss results obtained with these complementary
approaches in the following.

\subsubsection{Dynamic checking}
\label{sec:rqC-1}

To find dynamic evidence of parallelism, we ran the test suites from
our set of \numMedLong{} projects to output all key-value pairs of
Maven parameters.  To that end, we used the option~\CodeIn{-X} to
produce debug output and the option~\CodeIn{-DskipTests} to skip
execution of tests.  We skipped execution of tests as we observed from
sampling that only bootstrapping the Maven process suffices to infer
which parallel configuration modes it uses to run the
tests.  It is also important to point that we used the default
configurations specified in the project.  We inferred parallel
configurations by searching for certain configuration parameters in
log files. According to Maven's
documentation~\cite{maven-surefire-plugin}, a parallel configuration
depends either on (1) the parameter \CodeIn{parallel} to define the
parallelism mode within a JVM followed by the parameter
\CodeIn{threadCount} or (2) the parameter
\CodeIn{forkCount}\footnote{This parameter is named \CodeIn{forkMode}
  in old versions of Maven Surefire.} to define the number of forked
JVMs.  As such, we captured, for each project, all related key-value
pairs of Maven parameters and mapped those pairs to one of the
possible parallelization modes.  For instance, if a given project
contains a module with the parameter
\CodeIn{<forkCount>1C</forkCount>}, the possible classifications are
\ForkSeq{} or \ForkParMeth{}, depending on the presence and the value
of the parameter \CodeIn{parallel}.  If the parameter
\CodeIn{parallel} is set to \CodeIn{methods} the detected mode will be
\ForkParMeth{}.  Large projects may contain several test suites
distributed on different Maven modules potentially using different
configurations.  For those cases, we collected the Maven output from
each module discarding duplicates as to avoid inflating counts for
configuration modes that appear in several modules of the same
project. For instance, if a project contains two modules using the
same configuration, we counted only one occurrence.
Considering
our set of \numMedLong{} projects, we found that only
\numProjectsPar{} of those projects had parallelism enabled
by default, with only configurations \ParClassSeqMeth{},
\ParClassParMeth{}, and \ForkSeq{} being used. Configurations
\ParClassParMeth{} and \ForkSeq{} were the most popular among these
cases. Note that these results under-approximate real usage of
parallelism as we used default parameters in our scripts to spawn the
Maven process.  That decision could prevent execution of particular
test modules. Table~\ref{tab:freqmodes-dynamic} shows the
\numProjectsPar{} projects we identified where parallelism is enabled by default in Maven.

Column ``\emph{Subject}'' indicates the name of the project, column
\begin{wraptable}{r}{5cm}
\caption{Subjects with parallel test execution enabled by default.}
\label{tab:freqmodes-dynamic}
\footnotesize
\centering
\setlength{\tabcolsep}{2.5pt}
\begin{tabular}{llcr}
\toprule
\multirow{2}{*}{\emph{Group}} & \multirow{2}{*}{\emph{Subject}} & \emph{\# of} & \multirow{2}{*}{\emph{Mode}}\\%
   & & \emph{modules} &\\%
\midrule%
Medium & Californium & 2/20 & C2\\%
Medium & Chaos & 1/1 & C2\\%
Long & \Comment{apache} Flink & 66/74 & FC0\\%
Long & \Comment{apache logging-}Log4J2 & 25/28 & FC0\\%
Long & \Comment{javaslang }Javaslang & 3/3 & C3\\%
Medium & Jcabi \Comment{jcabi-github} & 1/1 & C3\\%
Long & \Comment{hazelcast hazelcast-}Jet & 6/7 & FC0\\%
Long & \Comment{apache} Mahout & 8/9 & FC0\\%
Long & \Comment{jankotek} MapDB & 1/1 & C3\\%
Medium & \Comment{apache} OpenNLP & 4/4 & FC0\\%
Medium & \Comment{yegor256} Rultor & 1/1 & C3\\%
Medium & \Comment{yegor256} Takes & 1/1 & C3\\%
Long & \Comment{vavr-io} Vavr & 3/3 & C3\\%
\bottomrule%
\end{tabular}
\end{wraptable}
``\emph{\# of modules}'' indicates the fraction of modules containing
tests that use the configuration of parallelism mentioned in column
``\emph{Mode}''.
We note that, considering these projects, the modules that do not use
the configuration cited use the sequential configuration \Seq{}.
For example, three modules (=28-25) from Log4J2 use sequential
configuration. It came as a surprise the observation that
no project used distinct configurations in their modules.

\subsubsection{Static checking}
\label{sec:rqC-2}
Given that the dynamic approach cannot detect parallelism manifested
through the default
configuration of projects, we also searched for indications of parallelism in build
files\Comment{ in the same sample set of \numMedLong{} projects}.  We
parsed all \emph{pom.xml} files under the project's directory and used
the same approach as in our previous analysis to classify
configurations.  We noticed initially that our approach was unable to
infer the configuration mode for cases where the decision depends on
the input (\eg,
\CodeIn{<parallel>\$\{parallel.type\}</parallel>}). For these
projects, the tester needs to provide additional parameters in the
command line to enable parallelization (\eg, \CodeIn{mvn test
  -Dparallel.type=classesAndMethods}). To handle those cases, we
considered all possible values for the parameter (in this case,
\CodeIn{\$\{parallel.type\}}).  It is also important to note that this
approach is not immune to false negatives, which can occur when
\emph{pom.xml} files are encapsulated in jar files or files downloaded from
the network.  Consequently, this approach complements the
the dynamic approach. Overall, we found \numProjectsParStatic{}
projects manifesting parallelism with this approach.
Compared to the set of projects listed in
Table~\ref{tab:freqmodes-dynamic}, we found four new projects, namely:
\CodeIn{Google Cloud\Comment{ Platform} DataflowJavaSDK} (using
configuration \ParClassParMeth), \CodeIn{Mapstruct} (using configuration
\ForkSeq{}), \CodeIn{T-SNE-Java} (using configuration \ForkSeq), and
\CodeIn{Urbanairship Datacube} (using configuration \ParClassParMeth).
Curiously, we also found that project \CodeIn{Jcabi}, \CodeIn{Rultor},
and \CodeIn{Takes} were not detected using this methodology.
That happened because these projects loaded a \emph{pom.xml} file from
a jar file that we missed.
Considering the static and dynamic methods together, we found a total
of \numProjectsParTotal{} distinct projects using parallelism,
corresponding to the union of the two subject sets.

\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQAdoptionOne{}:}~\emph{Results indicate that test
  suite parallelization is underused.  Overall, only
  \percentParallel{} of costly projects (\numProjectsParTotal{} out of \numMedLong)
  use parallelism.}
\end{minipage}
}
\end{center}

\begin{itemize}
  \item \numRQAdoptionTwo{}. \textbf{\RQAdoptionTwo{}}
\end{itemize}

To answer this research question we surveyed developers involved in a
selection of projects from our benchmark with time-consuming test
suites.  The goal of the survey is to better comprehend developer's
attitude towards the use of parallelism as a mechanism to speedup
regression testing.  We surveyed developers from a total of
\emailsProjects{} projects.  From the initial list of \numMedLong{}
project, we discarded 11 projects that we knew a priori used
parallelization, and \discartedProjects{} projects that we could not find
developer's emails from commit logs.  From this list of projects, we
mined potential participants for our study.  More precisely, we
searched for developer's name and email from the last 20 commits to
the corresponding project repository.  Using this approach, we
identified a total of \emailsSent{} eligible participants.  Finally,
we sent plain-text e-mails, containing the survey, to those developers.  In
total, \emailsAnswered{} developers replied but we discarded
\emailsFalseAnswers{} replies with subjective answers.  Considering
projects covered by the answers, a total of \emailsProjectsAnswered{}
projects (\percEmailsProjectsAnswered{} of the total) were represented
in those replies. Note that multiple developers on each project
received emails. In one specific case, one developer worked in
  multiple projects, and we consider it as a different answer. We sent the following set of questions to
developers:

\begin{enumerate}
\item How long does it take for tests to run in your environment? Can
  you briefly define your setup?
\item Do you confirm that your regression test suite does *not* run in parallel?
\item\label{questionThree} Select a reason for not using parallelization:
  \begin{enumerate}[label=\alph*)]
  \item I did not know it was possible
  \item I was concerned with concurrency issues
  \item I use a continuous integration server
  \item Some other reason. Please elaborate.
  \end{enumerate}
\end{enumerate}

Considering question 1, we confirmed that execution time was
compatible with the results we reported in Section~\ref{sec:rqA}.
Furthermore, \emailsCI{} of the participants indicated the use of
Continuous Integration (CI) to run tests, with \emailsDistributed{} of
these participants reporting that test suites are modularized and
those modules are tested independently in CI servers through different
parameters.  Those participants explained that such practice helps to
reduce time to observe test failures, which is the goal of speeding up
regression testing.  A total of \emailsLocal{} participants answered
that they do run tests in their local machines.  Note, however, that
CI does not preclude low-level parallelization.  For example,
installations of open-source CI tools (\eg{}, Jenkins~\cite{jenkins})
in dedicated servers would benefit from running tests faster through
low-level test suite parallelization.

Considering question 2, the answers we collected indicated, to our
surprise, that six of the \emailsProjectsAnswered{} projects execute
tests in parallel.  This mismatch is justified by cases where neither
of our checks (static or dynamic) could detect presence of
parallelism.  A closer look at these projects revealed that one of
them contained a \emph{pom.xml} file encapsulated in a jar file
(similar case as reported in Section~\ref{sec:rqC-2}), in one of the
projects the participant considered that distributed CI was a form of
parallelism, and in four projects the team preferred to implement
parallelization instead of using existing features from the testing
framework and the build system~---~in two projects the team
implemented concurrency control with custom JUnit test runners and in
two other projects the team implemented concurrency within test
methods.  Note that, considering these four extra cases (ignored two
distributed CI cases), the usage of parallelization increases from
\percentParallel{} to \percentParallelUpdated{}.  We do not consider
this change significant enough to modify our conclusion about
practical adoption of parallelization (\numRQAdoptionOne{}).

Considering question 3, the distribution of answers was as follows.  A
total of \emailsA{} of the \emailsProjectsAnswered{} developers who
answered the survey did not know that parallelism was available in
Maven (option ``a''), \emailsB{} of developers mentioned that they did
not use parallelism concerned with possible concurrency issues (option
``b''), \emailsD{} of developers mentioned that continuous integration
suffices to provide timely feedback while running only smoke
\begin{wrapfigure}{r}{0.15\textwidth}
    \centering
    \includegraphics[width=.15\textwidth]{results/survey.pdf}
    \caption{\label{fig:rq5-answers}Summary of developer's answers to
      survey question~\ref{questionThree}.}
\end{wrapfigure}
tests (\ie{}, short-running tests) locally (option ``c'')\Comment{here
  I want to say that they use it for something like "non-blocking
  testing" while developing in a local machine}, and \emailsD{} of
developers who provided an alternative answer (option ``d'') mentioned
that using parallelism was not worth the effort of preparing the test
suites to take advantage of available processing power.  A total of
\emailsNA{} of participants did not answer the last question of the
survey.  The pie chart in Figure~\ref{fig:rq5-answers} 
summarizes the distribution of answers.

\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQAdoptionTwo{}:}~\emph{Results suggest that dealing
    with concurrency issues (\ie{}, the extra work to organize test
    suite to safely explore concurrency) was the principal reason
    for developers not investing in parallelism.  Other reasons
    included availability of continuous integration services and
    unfamiliarity with the technology.}
\end{minipage}
}
\end{center}

\subsection{Speedups}
\label{sec:rqD}

\begin{itemize}
    \item \numRQSpeedupOne{}. \textbf{\RQSpeedupOne}
\end{itemize}

To answer \numRQSpeedupOne{}, we considered the \numProjectsPar{}
subjects from our benchmark that use parallelization \emph{by default}
(see Table~\ref{tab:freqmodes-dynamic}).  We compared running times
of test suites with enabled parallelization, as configured by project
developers, and without parallelization. It is important to note that
there are no observed failures in either execution.
Table~\ref{tab:speedup} summarizes results.
Lines are sorted by project names.
Columns ``\emph{Group}'' and ``\emph{Subject}'' indicate, respectively,
the cost group and the name of the project.
Column ``$T_s$'' shows sequential execution time and column ``$T_p$''
shows parallel execution time.
Column ``$T_s/T_p$'' shows speedup or slowdown.
As usual, a ratio above 1x indicates speedup and a ratio below 1x
indicates slowdown.

\begin{table}[t!]
\centering
\caption{
\label{tab:speedup}
Speedup (or slowdown) of parallel execution ($T_p$)
over sequential execution ($T_s$).  Default parallel configuration of
Maven is used.  Highest slowdown/speedup appears in gray color.}
%\resizebox{.41\textwidth}{!}{%
%  \scriptsize
\begin{tabular}{llrrr}
\toprule
\emph{Group} & \emph{Subject} & \multicolumn{1}{c}{$T_s$} & \multicolumn{1}{c}{$T_p$} & $T_s/T_p$ \\%
\midrule%
Medium & Californium & 1.45m & 1.40m & \cellcolor{lightgray}1.04x\\%
Medium & \Comment{BounceStorage} Chaos\Comment{ HTTP Proxy} & 1.51m & 1.47m & \cellcolor{lightgray}1.03x\\%
Medium &\Comment{ Apache }Flink& 11.79m & 2.57m & 4.59x\\%
Long &\Comment{ Apache }Log4J2& 8.24m & 8.21m & \cellcolor{lightgray}1.00x\\%
Medium &Javaslang& 2.18m & 1.82m & 1.20x\\%
Medium &Jcabi\Comment{ GitHub} & 2.76m & 0.30m & 9.20x\\%
Long &\Comment{ Hazelcast }Jet& 8.26m & 3.67m & 2.25x\\%
Long & \Comment{apache} Mahout & 27.38m & 18.15m & 1.51x\\%
Long &\Comment{ Jankotek }MapDB& 10.06m & 8.58m & 1.17x\\%
Medium & \Comment{apache} OpenNLP & 1.30m & 0.55m & 2.36x\\%
Medium & \Comment{yegor256} Rultor & 2.30m & 0.27m & 8.52x\\%
Medium & \Comment{yegor256} Takes & 2.00m & 0.19m & \cellcolor{lightgray}10.53x\\%
Long & \Comment{vavr-io} Vavr & 3.26m & 2.25m & 1.45x\\%
\midrule
Average &  &  &  & \avgSpeedup{}x\\
\bottomrule%
\end{tabular}
\end{table}

Results show that, on average, parallel execution was
\avgSpeedup{} times faster compared to sequential execution.
Three cases worth special attention: \CodeIn{Log4J2}, \CodeIn{Chaos},
and \CodeIn{Takes}.
We note that parallel execution in \CodeIn{Log4J2} was
ineffective.  We found that Maven invokes several test modules in this
project but the test modules that dominate execution time run
sequentially by default. This was also the case for the highlighted
project \CodeIn{Californium}.
No significant speedup was observed in \CodeIn{Chaos}, a project with
only three test classes, of which one monopolizes the bulk of test
execution time.
This project uses configuration \ParClassSeqMeth{}, which runs test
classes in parallel but runs test methods, declared in each class,
sequentially.
Consequently, speedup cannot be obtained as the cost of the single
expensive test class cannot be broken down with the selected
configuration.
Finally, the speedup observed in project \CodeIn{Takes} was
the highest amongst all projects. This subject uses configuration
\ParClassParMeth{} and contains 419 test methods distributed nearly
equally among 148 test classes with a small number of test methods.
Furthermore, several methods in those classes are time-consuming.
As result, the CPUs available for testing are kept occupied for the
most part during test execution.

\begin{center}
\fbox{
  \begin{minipage}{8cm}
    \textit{Answering \numRQSpeedupOne{}:}~\emph{Considering the
      machine setup we used, the average speedup observed with default
      configurations of parallelization was \avgSpeedup{}x.
      }
  \end{minipage}
}
\end{center}

\begin{itemize}
    \item \numRQSpeedupTwo{}. \textbf{\RQSpeedupTwo}
\end{itemize}

\newcommand{\subjectScalability}{MapDB}

This experiment evaluates the impact of making a growing number of
CPUs available to the build system for testing.  For this reason, we
used a different machine, with more cores, compared to the one described in
Section~\ref{sec:setup}.  We used a Xeon E5-2660v2 (2.20GHz) Intel
processor machine with 80 virtual CPUs (40 cores with two native
threads each) and 256GB of memory, running Ubuntu 14.04 LTS Trusty
Tahr (64-bit version). This experiment
spawns a growing number of JVMs in different CPUs, using parallel
configuration \emph{\ForkSeq{}}. We selected
subject \subjectScalability{} in this experiment as it represents the
case of a long-running test suite (see Table~\ref{tab:speedup}) with
test cases distributed across many test classes~--~194.  
Recall that a test class is the smallest unit that can be used to spawn a test job
on a JVM and that we have no control over which test classes will be
assigned to which JVM that the build system forks.
Figure~\ref{fig:scalability} shows the reduction in running times as
more CPUs contribute to the execution.
We ran this experiment for a growing number of cores 1, 3, ..., 39. 
The plot omits results beyond 17 cores as the tendency for higher
values is clear.
We noticed that improvements are marginal after three cores, which is
\begin{wrapfigure}{r}{0.23\textwidth}
  \includegraphics[width=0.23\textwidth]{R/scalability/scalability.pdf}
  \caption{\label{fig:scalability}Scalability.}
\end{wrapfigure}
the basic setup we used in other experiments.
This saturation is justified by the presence of a single test class,
\CodeIn{org.mapdb.WALTruncat}, containing 15 test cases that take over
two minutes to run.

\begin{center}
\fbox{
  \begin{minipage}{8cm}
    \textit{Answering \numRQSpeedupTwo{}:}~\emph{Results suggest that
      execution FC0 scales with additional cores but there is a bound
      on the speedup that one can get related to how well the test suite is
      balanced across test classes.}
  \end{minipage}
}
\end{center}

\subsection{Tradeoffs}
\label{sec:rq6-tradeoffs}

This dimension assesses the impact of using distinct parallel
configurations on test flakiness and speedup.  Increased parallelism
can increase resource contention leading to concurrency issues such as
data races across dependent
tests~\cite{luo-etal-fse2014,bell-etal-esecfse2015}.  Flakiness and
speedup are contradictory forces that could influence the decision of
practitioners about which parallel configuration should be used for
testing.  Note that Section~\ref{sec:rqD} evaluated speedup in
isolation.

\begin{itemize}
  \item \numRQIssuesOne{}. \textbf{\RQIssuesOne{}}
\end{itemize}

\input{flakiness-speedup-table}

To answer this research question, we selected 15 different subjects,
ran their test suites against all configurations described in
Section~\ref{sec:modes}, and compared their running times and rate of
test flakiness.  We used the sequential execution configuration,
\emph{\Seq{}}, as the comparison baseline in this experiment.  To
select subjects, we sorted projects whose test suites run in 1m or
more by decreasing order of execution time and selected the first
fifteen projects that use JUnit 4.7 or later.  The rationale for this
criteria is to ensure compatibility with parallel configuration since
older versions of JUnit does not support parallel testing.  We ran
each project on each configuration for \SubjectsReruns{} times.
Overall, we needed to reran test suites 270 times, 18 times (3x6
configurations) on each project.  Given the low standard deviations
observed in our measurements\Comment{ and the aggregated time cost of 270 test
executions}, we considered \SubjectsReruns{} reruns reasonable for this
experiment.

It is worth mentioning that we used custom JUnit runners as opposed to
Maven to run the test suites with different parallel configurations
(see Section~\ref{sec:modes}).  After carefully checking library
versions for compatibility issues and comparing results with JUnit's
we observed that several of Maven's executions exposed problems.  For
example, Maven incorrectly counts the number of test cases executed
for some of the cases where test flakiness are observed. These issues are
categorized and documented on our website~\cite{ourwebpage} and can be
reproduced with our scripts.  To address those issues we implemented
custom test runners for configurations \emph{\SeqClassParMeth{}},
\emph{\ParClassSeqMeth{}}, and \emph{\ParClassParMeth{}} and,  for configurations
\emph{\ForkSeq{}} and \emph{\ForkParMeth{}}, we
implemented a bash script that coordinates the creation of JVMs and
invokes corresponding custom runners.  As to faithfully reflect
Maven's behavior in our scripts, we carefully analyzed the source
code~\cite{maven-surefire-source} of the Maven Surefire plugin. We
implemented test runners using the \CodeIn{ParallelComputer} class from
JUnit~\cite{junit-parallel}.

We used Maven log files to identify test classes to run and used the
Maven dependency plugin~\cite{maven-dep} to build the project's
classpath (with the command \CodeIn{mvn dependency:build-classpath}).
Once we find the tests suite to run and the corresponding classpath,
we invoke the test runners mentioned above on them.  We configured
this experiment to run at most three JVMs in parallel.  Recall that in
our setup (see Section~\ref{sec:setup}), we limited our kernel to use
only three cores and reserved one core for OS-related processes.  To
ensure that our experiments terminate (recall that deadlock or
livelock could occur) we used the \CodeIn{timeout}
command~\cite{timeout-cmd} configured to dispatch a \emph{kill} signal
if test execution exceeds a given time limit. Finally, we save each
execution log and stack traces generated from JUnit to collect the
execution time, the number of failing tests, and to diagnose outliers
in our results.

\sloppy Table~\ref{tab:rq6-table} summarizes results ordered by
subject's name.  Values are averaged across multiple executions.  We
did not report standard deviations as they are very small in all
cases.  As to identify the potential causes of flakiness, we inspected
the exceptions reported in execution logs. We found that, in most
of the cases, flakiness was caused by race conditions: approximately
97.5\% of the failures were caused by a \CodeInF{null} dereference and 1.6\% were
caused by concurrent access on unsynchronized data structures.\Comment{
  More precisely, the manifested exceptions were
  \CodeInF{ConcurrentModificationException} (0.8\%),
  \CodeInF{NoSuchElementException} (0.7\%), and
  \CodeInF{ArrayIndexOutOfBoundsException} (0.7\%).}  Cases of likely
broken test dependencies were not as prevalent as race conditions
(0.8\% of the total): \CodeInF{EOFException} (0.2\%),
\CodeInF{FileSystemAlreadyExistsException} (0.2\%), and
\CodeInF{BufferOverflowException} (0.4\%). Results suggest that
anticipating race conditions to schedule test executions would have
higher impact compared to breaking test dependencies using a tool such
as ElectricTest~\cite{bell-etal-esecfse2015}.

The projects with flakiness
in all configurations were \CodeIn{AWS SDK}, \CodeIn{GoogleCloud}, and
\CodeIn{Moquette}.  It is worth highlighting the unfortunate case of
\CodeIn{Moquette}, which manifested more than 20\% flaky tests in
every configuration.  Considering time, it is noticeable from the
averages, perhaps as expected, an increasing speedup from
configuration \emph{\SeqClassParMeth} to \emph{\ParClassParMeth} and
from configuration \emph{\ForkSeq} to \emph{\ForkParMeth}.  It is also
worth mentioning that some combinations manifested slowdown instead of
speedup.  Recall that parallel execution introduces the overhead of
spawning and managing JVMs and threads.  Overall, results show that
0\% of flakiness have been reported in 30 of the 75 (=5x15)
pairs of project and configuration we analyzed (40\% of the total).
In for 4 of the 15 projects flakiness was not manifested in 
any combination pairs.  We noticed with some surprise that the average speedup of
configuration \emph{\SeqClassParMeth} was higher compared to
\emph{\ForkParMeth} indicating that it is not always the case that
using more CPUs pays off. Important to note that the cost of spawning
new JVMs can be significant in \emph{\ForkParMeth}.

\begin{center}
\fbox{
\begin{minipage}{8cm}
\textit{Answering \numRQIssuesOne{}:~Overall results indicate that the
  test suites of 73.33\% of the projects we analyzed could be run in
  parallel without manifesting any flaky tests.  In some of these
  cases, speedups were significant, ranging from 1x to 28.8x.}
\end{minipage}
}
\end{center}

%%  LocalWords:  RQ occurence parallelization Tradeoffs API readme th
%%  LocalWords:  mvn clearcut escapeinside xleftmargin untestable LTS
%%  LocalWords:  framexleftmargin CPUs Tahr sysstat gh Vagrantfile FC
%%  LocalWords:  javadoc isolcpus JUnit's JUnitCore Gligoric boxplots
%%  LocalWords:  outliers apache uber chaperone facebookarchive AWS
%%  LocalWords:  linkbench priori SDK GoogleCloud Moquette JVMs dir
%%  LocalWords:  xml basicstyle boxpos numberstyle deletekeywords JVM
%%  LocalWords:  uncompilable leftmargin quartile DskipTests forkMode
%%  LocalWords:  threadCount forkCount lrr BounceStorage Flink Gerrit
%%  LocalWords:  JenkinsCI Spotify Javaslang Jcabi Github Hazelcast
%%  LocalWords:  Jankotek MapDB Dparallel classesAndMethods Mapstruct
%%  LocalWords:  DataflowJavaSDK modularized JUnit llrrr lightgray rr
%%  LocalWords:  GitHub amongst Xeon runtime Scalability Dataflow sdk
%%  LocalWords:  JCTools RipMe classpath ParallelComputer livelock io
%%  LocalWords:  mapdb workflow XPath boxplot llcr javaslang jcabi
%%  LocalWords:  github hazelcast jankotek OpenNLP yegor Rultor vavr
%%  LocalWords:  Vavr SNE Urbanairship Datacube dereference
%%  LocalWords:  unsynchronized ConcurrentModificationException
%%  LocalWords:  NoSuchElementException EOFException ElectricTest
%%  LocalWords:  ArrayIndexOutOfBoundsException
%%  LocalWords:  FileSystemAlreadyExistsException
%%  LocalWords:  BufferOverflowException