### 2017-11-21 ase17/conclusions.tex
Renamed dir of ASE'17 paper to a proper name

\section{Conclusions}

Testing is expensive.  Despite all advances in regression testing
research, dealing with high testing costs remains an important problem
in Software Engineering.  This paper reports our findings on the usage
and impact of test execution parallelization in open-source projects.
Multicore CPUs are widely available today.  Testing frameworks and
build systems that capitalize on these machines also became popular.
Despite some resistance observed from practitioners, our results
suggest that parallelization can be used in many cases without
sacrificing reliability. More research needs to be done to improve
automation (\eg{}, breaking test dependencies, refactoring test
suites, enforcing safe test schedules) as to safely optimize parallel
execution.  The artifacts we produced as result of this study are
available from the following web page:

\begin{center}
  \webpage{}
\end{center}

%% Overall, this study brings to light the benefits and burdens of test
%% suite parallelization to improve test efficiency. It provides
%% recommendations to practicioners and developers of new techniques and
%% tools aiming to speed up test execution with parallelization.

%% Considering a set of \numSubjs{} popular Java projects we analyzed, we
%% found that \percentMedLongRunning{} of the projects contain costly
%% test suites but parallelization features still seem underutilized in
%% practice~---~only \percentParallelUpdated{} of costly projects use
%% parallelization.  The main reported reason for adoption resistance was
%% the concern to deal with concurrency issues.  Results suggest that, on
%% average, developers prefer high predictability than high performance
%% in running tests.

%%  LocalWords:  parallelization Multicore CPUs refactoring