### 2017-12-10 journal/anatomy.tex
Added IEEE Journal formatting

\section{Anatomy of a \pef{} failure}

It is important to note that previous work investigated test flakiness
manifested in sequential executions\Fix{cite cite}.  For example, a
test case can fail because a timing assumption of an asynchronous test
was broken.  Treating this source of flakiness is important but
orthogonal to this work: parallel execution flakiness can arise even
without sequential execution flakiness.  In the following we explain
how \pef{}s come to be.

\Fix{...}

\subsection{Detecting flakiness.}
\label{sec:detect-flakiness-build}

A popular approach to detect test flakiness is to iteratively
re-execute (parts of) the test suite multiple times marking as flaky
those tests that produce both pass and failing outputs.  We refer to
this process as IRE (for Iterative RE-execution).

Considering the Maven\Fix{is this maven or maven's surefire
  plugin~\cite{maven-surefire-plugin}?} build system, the process
starts with the developer determining a bound N for the number of
re-executions of the test suite.  Only tests that fail in one given
iteration are scheduled for re-execution in the next iteration.  At
the end of the process, a test is considered passing if it passes in
the first iteration, a test is considered failing if it fails in all
iterations, and it is considered flaky if it fails in all iterations
but the last it executed.  The process terminates when all tests
eventually pass or when the bound N is reached.  Note that every
iteration runs a subset of the tests from the previous iteration so,
assuming most tests will pass in the first iterations, execution cost
is much lower compared to executing the entire test suite N times.
However, note also that this method can particularly miss flaky tests
for the following reasons: \Fix{elaborate: (1) passing tests can fail
  in subsequent iterations and (2) ignoring tests may prevent
  manifestations of read-write conflicts.}

\Fix{explain other method...}

%%  LocalWords:  iteratively