### 2017-12-10 journal/study.tex

\section{Empirical Study}

Table~\ref{table:cost}...\Fix{Why Forked JVM-L1 is (in some cases) so much better
compared to Forked JVM-L0? (io?)}\Fix{Why Forked JVM seems to be
  superior to all other Single JVM methods?  Are all CPUs in use?  If
  not, why one would use Single JVM (given that Forked JVM is faster
  and reduces flakiness)?}

Table~\ref{table:failures}...\Fix{I expected ``Single JVM + L3'' to
  find most flaky tests.}\Fix{Need to compute these results with
  another IRE method.}

\begin{table*}[t]
  \centering
  \begin{tabular}{|c|r|r|r|r|r|r|r|}
    \hline
    \multirow{2}{*}{\emph{Subject}} & \multirow{2}{*}{\emph{\# Tests}} &  \multicolumn{6}{c|}{\emph{Time}}\\
    \cline{3-8}
    & & \multicolumn{4}{c|}{\emph{Single JVM}} & \multicolumn{2}{c|}{\emph{Forked JVMs}} \\
%    & & \multicolumn{1}{c|}{\forkSeq{}} &
%    \multicolumn{1}{c|}{\forkDia{}} &
%    \multicolumn{1}{c|}{\singleJVMSeq{}} &
%    \multicolumn{1}{c|}{\singleJVMDia{}} &
%    \multicolumn{1}{c|}{\singleJVMClass{}} &
    %    \multicolumn{1}{c|}{\singleJVMMethod{}}\\
    \cline{3-8}
    & &  \multicolumn{1}{c|}{\Seq{}} &
    \multicolumn{1}{c|}{\ParClassSeqMeth{}} &
    \multicolumn{1}{c|}{\SeqClassParMeth{}} &
    \multicolumn{1}{c|}{\ParClassParMeth{}} &
    \multicolumn{1}{c|}{\Seq{}} &
    \multicolumn{1}{c|}{\ParClassSeqMeth{}}\\
    \hline
    \hline
    \multicolumn{8}{|c|}{\emph{Longer executions}} \\
    \hline
    camel-core & 5,683 & 24m:52s & 1.00x & 1.00x & 1.00x & 5.88x & 5.82x \\
    \hline
    jetty.project & 543 & 4m:14s & 2.41x & 2.97x & 5.61x & 2.44x & 13.70x \\
    \hline
    jgit & 3,422 & 3m:59s & 2.68x & 2.37x & 1.95x & 2.31x & 4.50x \\
    \hline
    okhttp & 1545 & 1m:16s & crash I & crash II & crash II & 3.40x & TIMEOUT \\
    \hline    
    \hline
    \multicolumn{8}{|c|}{\emph{Shorter executions}} \\
    \hline
    graphhopper & 1,259 & 12s & 0.92x & 1.01x & 0.53x & 1.72x & 1.17x \\
    \hline
    junit4 & 791 & 10s & 0m10.08s & 0m10.32s & 0m10.49s & 0m10.13s & 0m9.96s \\
    junit4** & <TESTCASES> & 10s & TIMEOUT & crash II & crash II & 0.99x & FIX \\
    \hline
    retrofit & 263 & 6s & 1.25x & 1.08x & 1.22x & 1.11x & 1.14x \\
    \hline
  \end{tabular}
  \caption{\label{table:cost}Test execution time.}
\end{table*}

\Fix{junit4 - need to look more carefully the log file}
\Fix{FIX - I used a wrong configuration in the pom.xml file}
\Fix{Camel - still couldn't find an explanation to unexpected behavior with parallel settings}
\Fix{Crash I - Surefire internal error (NullPointerException)}
\Fix{Crash II - JVM Crash (e.g. System.exit)}
\Fix{TIMEOUT - Possibly a deadlock: Execution could not finish within 30min}

\begin{table*}[t]
  \centering
  \begin{tabular}{|c|r|r|r|r|r|r|r|}
    \hline
    \multirow{2}{*}{\emph{Subject}} & \multirow{2}{*}{\emph{\# Tests}} &  \multicolumn{6}{c|}{\emph{Time}}\\
    \cline{3-8}
    & & \multicolumn{4}{c|}{\emph{Single JVM}} & \multicolumn{2}{c|}{\emph{Forked JVM}s} \\
    \cline{3-8}
    & & \multicolumn{1}{c|}{\Seq{}} &
    \multicolumn{1}{c|}{\ParClassSeqMeth{}} &
    \multicolumn{1}{c|}{\SeqClassParMeth{}} &
    \multicolumn{1}{c|}{\ParClassParMeth{}} &
    \multicolumn{1}{c|}{\Seq{}} &
    \multicolumn{1}{c|}{\ParClassSeqMeth{}} \\
    \hline
    \hline
    \multicolumn{8}{|c|}{\emph{Longer executions}} \\
    \hline
    camel (core) & 5,979 & 0 & ! & ! & ! & 75 & 70 \\
    \hline
    jetty (client) & 535 & 0 & 106** & 0 & 76** & 0 & 111 \\
    \hline
    jgit & 3,364 & 0 & 75 & 57* & 99* & 1 & 79 \\
    \hline
    okhttp & 1,525 & 0 & - & - & ! & 0 & ! \\    
    \hline
    \hline
    \multicolumn{8}{|c|}{\emph{Shorter executions}} \\
    \hline
    graphhopper & 1,259 & 0 & 152 & 55 & 347 & 14 & 67 \\
    \hline    
    junit4 & 791 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    \hline
    retrofit & 261 & 0 & 1 & 0 & 1 & 0 & 1 \\
    \hline
  \end{tabular}
  \caption{\label{table:failures}Number of flaky tests obtained with
    Maven's IRE (see Section~\ref{sec:detect-flakiness-build}) using
    bound of \Fix{?}.\Fix{* = Contains a skipped test. I don't know
      yet why a failing test may be skipped when it is executed
      individually}\Fix{Reminder: To not be confused with Sequential
      Flakiness. Sequential Flakiness must not be considered}}
\end{table*}