### 2019-06-02 tse19/main.tex
pass up to description of RQs

\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{graphicx}
% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\input{macros}

\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{balance}
\usepackage{float}
\usepackage{url}
\usepackage{tabularx}
\usepackage[table]{xcolor}
\usepackage{booktabs}
\usepackage{amsmath} 
\usepackage{subfigure} 
\usepackage{multirow}
\usepackage{mdframed}
\usepackage{xcolor}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Test Suite Parallelization in Open-Source Projects:\\ A Study on Its Usage and Impact}

\author{Sotero Junior, Jeanderson Candido, Marcio Ribeiro and~Marcelo d'Amorim

% <-this % stops a space
\IEEEcompsocitemizethanks{
\IEEEcompsocthanksitem 
S. Junior and M. d'Amorim are with the Universidade Federal de
Pernambuco, Recife, PE, Brazil.
E-mail: \{srsj2, damorim\}@cin.ufpe.br

\IEEEcompsocthanksitem 
J. Candido is with the Delft University of Technology, The Netherlands.
E-mail: jeandersonbc@gmail.com

\IEEEcompsocthanksitem 
M. Ribeiro is with the Computing Institute, Federal University of Alagoas, Maceio, AL 57072-900, Brazil.
E-mail: marcio@ic.ufal.br
}% <-this % stops an unwanted space
\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~45, No.~6, June~2019}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}

\IEEEtitleabstractindextext{%
\begin{abstract}
Dealing with high costs of testing remains an important problem in
Software Engineering.  Test suite parallelization is an important
approach to address this problem. \Mar{Sotero, vc. atualizou estas
  macros?} \textcolor{red}{[S.J: atualizado.]} Considering a set of \numSubjs{} Java projects we analyzed
from \github\, we found that \percentMedLongRunning{} of the them
contain long-running test suites, but parallelization features still
seem underutilized in practice~---~only \percentParallelUpdated{} of
projects with long-running test suites parallelize test
execution. According to a survey we conducted, the main reason for
adoption resistance was the concern to deal with concurrency issues
that could result in test flakiness.
%% This Results suggest that, on
%% average, developers prefer high predictability than high performance
%% in running tests.

\Fix{revise...}
This paper reports our findings on
the usage and impact of test suite parallelization in open-source
projects. It provides recommendations to practitioners and
tool developers to speed up test execution. A tool is available to \Fix{...}.

\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
test suite, parallelization, Maven, Java, Software Engineering.
\end{IEEEkeywords}}

% make the title area
\maketitle
\IEEEdisplaynontitleabstractindextext
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{
\section{Introduction}\label{sec:introduction}}

\IEEEPARstart{D}{ealing} with high testing costs has been an important
problem in software engineering research and industrial practice.
Several approaches have been proposed in the literature to address
this problem, with the focus mainly on test suite minimization,
prioritization, reduction, and selection~\cite{yoo-harman-stvr2012}.
In industry, the focus has been mainly on distributing the testing
workload.  Evidence of this are the Google TAP
system~\cite{google-tap,google-ci} and the Microsoft CloudBuild
system~\cite{prasad-shulte-ieee-microsoft-ci}, which provide
distributed infrastructures to efficiently build massive amounts of
code and run tests.

%Distributing testing workloads in-house is also a
%viable solution at smaller-size organizations.

At large organizations, the alternative of renting cloud
services~\cite{cloud-services} or even building proprietary
infrastructures for building and running tests is a legitimate
approach to mitigate the regression testing problem.  For example, as
of August 2013, the test suite of the Groupon PWA system, which powers
the \url{groupon.com} website, included over 19K tests.  To run all
those tests under 10m, Groupon used a cluster of 4 computers with 24
cores each~\cite{kim-etal-fse2013}. However, for projects with modest
or nonexistent budgets and yet relatively heavy testing workloads,
this solution may not be economically viable.  For these cases, the
use of commodity hardware\Comment{ (\eg{}, existing workstations)} is
an attractive solution for running tests.  The proliferation of
multi-core CPUs and the increasing popularization of testing
frameworks and build systems, which today provide mature support for
parallelization, enable speedups through
increased CPU usage~(see Section~\ref{sec:background}).  These two
elements~---~demand for cost-effective test execution and supply of
relatively inexpensive testing infrastructures~---~inspired us to
investigate test suite parallelization in practice.

This paper reports on an empirical study we conducted to analyze the
usage and impact of low-level parallelization to speed up testing in
open-source projects.  This is a relevant problem given the tremendous
popularity of open-source development and regression testing
research~\cite{yoo-harman-stvr2012}.  Note that parallelization is
complementary to other approaches to mitigate testing costs such as
(safe) test
selection~\cite{Rothermel:1997:SER:248233.248262,gligoric-etal-issta2015}
and continuous
integration~\cite{Saff:2003:RWD:951952.952340}.\Mar{revisit this later
  on...} In addition, we propose a solution to unleash the power of
test parallelization reducing (\Fix{eliminating?}) test flakiness due
to race conditions.

This study covers the following dimensions of interest
(i)~feasibility, (ii) adoption, (iii) speedup, and (iv) tradeoffs.
The dimension \emph{feasibility} measures the potential of
parallelization to reduce testing costs.  In the limit,
parallelization would be fruitless if all projects had short-running
test suites or if the execution cost was dominated by a single test
case in the suite.  The dimension \emph{adoption} evaluates how often
existing open-source projects use parallelization schemes and how
developers involved in costly projects---\ie{}, project not using test suite
parallelization---perceive this technology.  It is important to evaluate
resistance of practitioners to the technology and to understand their
reasons.  The \emph{speedup} dimension evaluates the observed impact
of parallelization in running times on projects that already use it.  Finally, the dimension
\emph{tradeoffs} evaluates the relationship between speedups obtained
with parallelization and issues that arise when running tests in
parallel, including test
flakiness~\cite{luo-etal-fse2014,bell-etal-esecfse2015}.  We briefly
summarize our findings in the following.

\Mar{We need to reorganize this a bit as to accomodate the new
  addition to the paper...Revisit this later...}

\noindent\emph{Feasibility.}~To assess how prevalent long-running test suites
are we selected \numSubjs{} popular Java projects from \github{} containing
Maven build files~\cite{maven}.  Section~\ref{sec:eval} details our methodology
to select subjects and to isolate our experiments from environmental noise.
Results indicate that nearly \percentMedLongRunning{} of the projects take at
least 1m to run and \percentLongRunning{} of the projects take at least 5m to
run.  Considering the \numMedLong{} projects with test suites taking longer than
a minute to run, the average execution time of a test suite was
\averageMedLongRunning{}.  Results also show that test cases are typically
short-running, typically taking less than half a second to run.  Furthermore, we
found that only in rare cases few test cases monopolize the overall time to run
a test suite.

\noindent\emph{Adoption.}~ We considered two aspects in measuring technology
adoption.  First, we measured usage of parallelism in open-source projects.
Then, we ran a survey with developers to understand the reasons that could
explain resistance to using the technology.\Comment{ For the quantitative
analysis we checked both statically (parsing build files) and dynamically
(monitoring execution of build files) for the presence of parallelization.}
Considering only the projects whose test suites take longer than a minute to
run, we found that only \percentParallelUpdated{} of them use parallelism.  We
also contacted developers from a selection of costly projects that did not use
parallelization to understand the reasons for not using parallelization.\textcolor{red}{[S.J: nao fizemos isso.]}
Dealing with concurrency-related issues (\eg{}, the extra work to organize test
suite to avoid concurrency errors) and the availability of continuous
integration services were the most frequently answered reasons for not
considering parallelization.

\noindent\emph{Speedups.}~We used two setups to measure speedups.  In one setup
we measured speedups obtained on projects that run test suites in parallel by
default.  In the other setup, we evaluated how execution scales with the number
of available cores in the machine.\Comment{We did not change parallel
configurations on that setup.  For comparison, we disabled parallel execution to
obtain the running time of a sequential execution.}  Considering the first
setup, results indicate that the average speedup of parallelization was
\avgSpeedup{}x.  Although we found cases with very high speedups (\eg{}, 5.77x
for project Twilio Java), we also found cases where the speedups were not very
significant.  Considering the scalability experiment, we noticed, perhaps as
expected, that parallelization obtained with forking JVMs scales with the number
of cores but the speedups are bounded by long-running test classes.

\noindent\emph{Tradeoffs.}~Test flakiness is a central concern when running
tests in parallel.  Dependent tests can be affected by different schedulings of
test methods and classes.  This dimension of the study measures the impact of
different parallel configurations on test flakiness and speedup.  Overall,
results indicate that configurations that fork JVMs do not achieve speedups as
high as other more-aggressive configurations, but they manifest much lower
flakiness ratios.

Our observations may trigger different actions:

%\setlist[itemize]{leftmargin=1.1em}
\begin{itemize}
\item \emph{Incentivize forking.}~Forked JVMs manifest very low rates
  of test flakiness.  Developers of projects with long-running test
  suites should consider using that feature, which is available in
  modern build systems today (\eg{}, Maven).
\item \emph{Break test dependencies.}~Non-forked JVMs can achieve
  impressive speedups at the expense of sometimes impressive rates of
  flakiness.  Breaking test dependencies (with
  ElectricTest~\cite{bell-etal-esecfse2015}, for example) to avoid flakiness is advised for developers with
  greater interest in efficiency.
\item \emph{Refactor tests for load balancing.}~The configuration with
  forked JVMs scales
  better\Comment{ with the number of cores} when the test workload is balanced
  across testing classes.  Automated refactoring could help balance
  the workload in scenarios where developers are not willing to change
  test code but have access to machines with a high number of cores.
\item \emph{Improve debugging for build systems.}~While preparing our
  experiments, we found scenarios\Comment{, related to test
    parallelization,} where Maven's executions did not reflect
  corresponding JUnit's executions.\Comment{(Docker reproduction scripts
  available.)} Those issues can hinder developers from using parallel
  testing. Better debugging support for build systems could help
  on that. 
\end{itemize}

The artifacts we produced as result of this study are available from
the following web page \webpage{}.

%\hfill lorem ipsum

%\hfill August 26, 2015

% SECTION 2
\section{\label{sec:background}Parallel Execution of Test Suites}

\begin{wrapfigure}{r}{0.13\textwidth}  
  \vspace{-2mm}
  \centering
  %  \includegraphics[width=0.35\textwidth]{figs/parallel-levels.pdf}    
  \includegraphics[width=2cm]{figs/parallel-levels-short.pdf}  
  \caption{\label{fig:levels}Levels of parallelism.}
  \vspace{-2mm}
\end{wrapfigure}
This section explains basic concepts related to test parallelization
in Java projects. Figure~\ref{fig:levels} illustrates different levels
of parallelism in test execution.  The highest level indicates
parallelism obtained through different machines on the network. For
instance, using virtual machines from a cloud service to distribute
test workload. The Google TAP~\cite{google-tap,google-ci} and the
Microsoft CloudBuild~\cite{prasad-shulte-ieee-microsoft-ci} systems
are examples of this kind of parallelization. The lowest levels
illustrate parallelism obtained within a single machine. These levels
are complementary to each other: the lowest levels leverage the
computing power of individual nodes whereas the highest level
leverages the aggregate processing power of a network of
machines. This paper focuses on low-level parallelism, where
computation can be offloaded at different CPUs within a machine and at
different threads within each CPU.  This form of parallelism is
enabled through build systems---by spawning processes in different
CPUs---and testing frameworks---by spawning threads in one given CPU.
It is important to note that a variety of testing frameworks provide
today support for parallel test execution (e.g.,
JUnit~\cite{junit-org}, TestNG~\cite{testng}, and NUnit~\cite{nunit})
as to benefit from the available power of popular multi-core
processors.  In the following, we elaborate relevant features of
testing frameworks and build systems for parallelization.  We focused
on Java, Maven, and JUnit but the discussion can be generalized to
other language and tools.

\subsection{Testing Frameworks}
\label{sec:frameworks}

The list below shows the choices to control parallelism within one
Java Virtual Machine (JVM).  These options are offered by the testing
framework (\eg{}, JUnit).

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~No parallelism is involved.
\item
    \textbf{Sequential classes; parallel methods
      (\SeqClassParMeth).}~This configuration corresponds to running
    test classes sequentially, but running the test methods from those
    classes concurrently.
\item
    \textbf{Parallel classes; sequential methods
      (\ParClassSeqMeth{}).}~This configuration corresponds to running
    test classes concurrently, but running test methods sequentially.
\item
    \textbf{Parallel classes; Parallel methods
      (\ParClassParMeth).}~This configuration runs test classes and
    methods concurrently.\Comment{corresponds to the
    combination of \SeqClassParMeth{} and \ParClassSeqMeth{}.  }
\end{itemize}

An important aspect to consider when designing new test suites or when
deciding which configuration to use for existing test suites is the
possibility of race conditions on shared data during test
execution. Although conceptually test cases are independent, they
often share data\Fix{cite Mike's paper, bell-etal-esecfse2015}. Data
sharing can occur, for example, through state that is reachable from
statically-declared variables in the program or through variables
declared within the scope of the test class or even through resources
available on the file system and the network~\cite{luo-etal-fse2014}.

Test flakiness and execution speedups depend on a number of
factors. Considering flakiness, a configuration may be preferrable to
another depending on the characteristics of the project and its test
suite. For example, configuration \ParClassSeqMeth{} may be preferable
over \SeqClassParMeth{} when it is clear that several test methods in
a given class perform operations involving shared data. As
configuration \ParClassParMeth{} does not restrict scheduling
orderings, it is more likely to manifest data races during execution
if there are sources of data sharing across test classes and their
methods. Note also that speedups depend on several factors, including
the test suite size and the distribution of test methods per class.

\subsection{Build Systems}
\label{sec:builder}

%% The build system can spawn multiple JVMs, each running on its own OS
%% process on a given CPU and handling a partition of the test set.
Forking OS processes to run test jobs is the basic mechanism of build
systems to obtain parallelism at the machine space (see
Figure~\ref{fig:levels}).  For Java-based build systems, such as Maven
and Ant, this amounts to spawning one JVM, on a given CPU, to handle a
test job and aggregating results when jobs finish.  The list below
shows the choices to control parallelism through the build system
(\eg{}, Maven).

\begin{itemize}
\item
  \textbf{Forked JVMs with sequential methods (\ForkSeq).}~The build
  system spawns multiple JVMs with this configuration, assigning a
  partition of the set of test classes to each JVM.  Test classes and methods
  run sequentially within each JVM as \Seq\ does.
\item
  \textbf{Forked JVMs with parallel methods (\ForkParMeth).}~As
  \ForkSeq{}, the build system forks multiple JVMs with this
  configuration, but it runs test methods concurrently, as
  \SeqClassParMeth{}.
\end{itemize}

%% In the configuration \ForkSeq{}, each spawned
%% process runs a different test class at time and the builder merges the
%% results from each execution. In the configuration \ForkParMeth{}, the
%% build system forwards settings to the underlying testing framework to
%% enable parallel execution within each process in addition to running
%% test classes on different processes.

Note from the listing that forking can only be combined with
configuration \SeqClassParMeth{} (see Section~\ref{sec:frameworks}) as
Maven made the design choice to only accept one test class at a time
per forked process. Intuitively, this option enables safe execution of
tests when dependent methods exist across different classes\Fix{cite
  Mike's work}. For the cases the developer prepared code against that
threat, Maven offers an option to ``reuse'' JVMs as to avoid the
non-negligible cost of creating JVM instances.

%% \Mar{$\leftarrow$Are you sure about this?
%%   It seems silly.  Maybe, the rationale is to ``keep design
%%   simple''.}\Jbc{I'm sure. in addition to keep design simple, I feel
%%   like this was initially conceived to run tests in isolation}
%% \Mar{$\leftarrow$confirma?  se sim, pode mostrar isto na
%%   configuracao abaixo?}\Jbc{added...}
%%We are unaware of other build system capable of running multiple
%%classes at time within a forked process.
%%  to
%% define tasks related to the project building and these tasks are
%% performed several plugins.This file
%% contains defintions of tasks, which are implemented by a collection of
%% plugins.
%% \Jbc{falar do surefire e explicar a
%% configuracao ilustrada - Surefire levanta uma JVM por core e cria uma
%% pool de 5 threads em cada JVM para executar testes em paralelo.}
\begin{figure}[h!]
\centering
\scriptsize
\lstset{
  escapeinside={@}{@},
  numbers=none,xleftmargin=1em,frame=none,framexleftmargin=0.5em,
  basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
  morekeywords={parallel, threadCount, perCoreThreadCount,
  forkCount, reuseForks},
  deletekeywords={true}
}
\vspace{-4ex}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <forkCount>1C</forkCount>
        <threadCount>5</threadCount>    
        <parallel>methods</parallel>
        <reuseForks>true</reuseForks>        
    </configuration>
</plugin>
\end{lstlisting}
  \vspace{-4ex}
  \caption{\label{fig:surefire} Fragment of a Maven pom.xml file with
    an \ForkParMeth{} configuration.}
\end{figure}

\noindent\emph{Example:} Figure~\ref{fig:surefire} shows a
fragment of a Maven configuration file, known as \pomf{}, highlighting
options to run tests using the parallel execution mode \ForkParMeth{}.
Maven implements this feature through its Surefire JUnit test
plugin~\cite{maven-surefire-plugin}.  With this configuration, Maven
forks one JVM per core (\CodeIn{forkCount} parameter) and uses five
threads (\CodeIn{threadCount} parameter) to run test methods
(\CodeIn{parallel} parameter) within each forked JVM. In addition, Maven reuses
created JVMs on subsequent forks when execution of a test class
terminates (\CodeIn{reuseFork} parameter). This is an example of a
very aggressive setup, \ie{}, a setup that could substantially speedup
execution but could introduce test flakiness if test code is not
prepared for parallel execution.

% SECTION 3

\section{Objects of Analysis}
\label{sec:subjects}

\begin{figure}[t]
%  \vspace{-5mm}
  \centering
  \includegraphics[width=0.37\textwidth]{results/piechart-subjs.pdf}
  \caption{\label{fig:subjects}Breakdown of project by sources of
    issues.}
\end{figure}

\label{sec:setup}

We used the Python library PyGitHub~\Fix{cite}, which uses the
\github{}'s search API~\cite{githubsearch}, to identify projects that
satisfy the following criteria: (1) the primary language is
Java\footnote{In case of projects in multiple languages, the \github{}
  API considers the predominant language as the primary language.},
(2) the project has at least 100 stars, and (3) the latest update was
on or after January 1st, 2016. Although there is no clearcut limit on
the number of \github{} stars~\cite{github-stars} to define relevant
projects, we observed that one hundred stars was enough to eliminate
trivial subjects. The third criteria serves to skip projects without
recent activity.

To sum up, after obtaining the list of potential projects from GitHub,
we filtered those containing a \pomf{} file in the root
directory.\Comment{ A Maven project may contain several sub-modules
  with multiple \pomf{} files.}  Then, considering this set of Maven
projects, we executed the test suite for \SubjectsReruns{} times
discarding those projects with issues. Figure~\ref{fig:subjects}
shows the breakdown of projects by source of issue. As of
March 2019, our search returned a total of \SubjectsGithub{}
subjects.  We found that \SubjectsGithubNotMaven{}
projects in this set were not Maven. Compared with our previous
experiment~\cite{Candido:2017:TSP:3155562.3155667}, executed nearly 2
years ago, we observed  that many more
projects now use Gradle---73\% of the \SubjectsGithubNotMaven{} non-Maven
projects use Gradle. Although parallelization can
be used in Gradle, we decided to keep using using Maven for
consistency with the previous experiment. From the remaining set of
projects, \SubjectsGithubNotTestable{} were discarded
because of environment incompatibility (\eg, missing\Comment{ required
  web browser or database management system}~DBMS) and
\SubjectsGithubFlaky{} projects were discarded because of flaky
tests~\cite{luo-etal-fse2014}. A flaky test is a test that passes or
fails non-deterministically resulting in wasted effort.  As some of our experiments consist of running tests on
different threads, we ignored these projects as it would be
impractical to identify whether a test failed due to a race condition
or some other source of flakiness. Note that existing techniques exist to
detect and mitigate flakiness~\Fix{cite cite}, so we considered this
problem out of scope. From the remaining
\SubjectsGithubConsistant{} projects with test execution
manifesting deterministic results, we
eliminated \SubjectsGithubTooManyFailures{} projects that contain
\SuiteFailingThreshold{} or more failing tests as to reduce
bias. For projects with less failing tests, we used the JUnit's
\CodeIn{@Ignore} annotation to discard the failing tests as we did not
want a premature termination of test run to influence in timings.  Our final set of
subjects include  \numSubjs{} projects.

 %% \textcolor{red}{[S.J: Foi possivel observar que houve uma consideravel conversao de ferramenta de automacao, do Maven para Gradle,  dos 3154 subject que nao eram Maven, 2302 eram Gralde (73\%). Projeto como o MapDB, que e bastante citado no paper de ase17, atualmente converteu-se a Gradle. ]}

% SECTION 4

\section{Evaluation}
\label{sec:eval}

We pose the following research questions, organized by the dimensions
of analysis we presented in Section~\ref{sec:introduction}.

%% Feasibility
\newcommand{\numRQFeasibilityOne}{RQ1}
\newcommand{\RQFeasibilityOne}{How prevalent are time-consuming
  test suites\Comment{ in open-source projects}?}

\newcommand{\numRQFeasibilityTwo}{RQ2}
\newcommand{\RQFeasibilityTwo}{How is time distributed across test cases?}

%% Adoption
\newcommand{\numRQAdoptionOne}{RQ3}
\newcommand{\RQAdoptionOne}{How popular is test suite
  parallelization\Comment{ in open-source projects}?}

\newcommand{\numRQAdoptionTwo}{RQ4}
\newcommand{\RQAdoptionTwo}{What are the main reasons that prevent developers
  from using test suite parallelization?}

%% Speedups
\newcommand{\numRQSpeedupOne}{RQ5}
\newcommand{\RQSpeedupOne}{What are the speedups obtained with parallelization
  (in projects that actually use it)?}

\newcommand{\numRQSpeedupTwo}{RQ6}
\newcommand{\RQSpeedupTwo}{How test execution scales with the number of
  available CPUs?}

%% Tradeoffs
\newcommand{\numRQIssuesOne}{RQ7}
\newcommand{\RQIssuesOne}{How parallel execution configurations affect testing
  costs and flakiness?}

\setlist[itemize]{leftmargin=1em}
\begin{itemize}
\item Feasibility
  \begin{itemize}
  \item \textbf{\numRQFeasibilityOne.} \RQFeasibilityOne
  \item \textbf{\numRQFeasibilityTwo.} \RQFeasibilityTwo    
  \end{itemize}
\item Adoption
  \begin{itemize}
  \item \textbf{\numRQAdoptionOne.} \RQAdoptionOne    
  \item \textbf{\numRQAdoptionTwo.} \RQAdoptionTwo
  \end{itemize}
\item Speedups
  \begin{itemize}
  \item \textbf{\numRQSpeedupOne.} \RQSpeedupOne
  \item \textbf{\numRQSpeedupTwo.} \RQSpeedupTwo
  \end{itemize}      
\item Tradeoffs
  \begin{itemize}
  \item \textbf{\numRQIssuesOne.} \RQIssuesOne    
  \end{itemize}
\end{itemize}

To run our experiments, we used an Intel Xeon ES-2650 v4 (2.20 GHz)
with eight virtual CPUs (four cores with two native threads each) and
16GB of memory, running Ubuntu 18.10 Cosmic Cuttlefish (64-bit version).  We
used\Comment{ Git,} Java 8 and Maven 3.3.9 to build projects and run test
suites. To process test results and generate plots we used Python\Comment{ 3.4},
Bash, R and Ruby\Comment{ 2.3}.  All source artifacts are publicly available for
replication on our website~\cite{ourwebpage}.  This includes supporting
scripts\Comment{ (\eg, the scripts to run the tests and generate raw analysis
data)} and the full list of projects.

\subsection{Feasibility}
\label{sec:rqA}
\label{sec:rqB}

\begin{itemize}
  \item \numRQFeasibilityOne{}. \textbf{\RQFeasibilityOne}
\end{itemize}

To evaluate prevalence of projects with time-consuming test suites, we
considered the \numSubjs{} projects, appearing in 
Figure~\ref{fig:subjects}.  Figure~\ref{fig:mvn-execution} illustrates
the script we used to measure time.

We took the following actions to isolate our environment from
measurement noise.
First, we observed that some test tasks called test-unrelated tasks
(\eg, \emph{javadoc} generation and static analyses) that could
interfere in our time measurements.
To address that potential issue, we inspected Maven execution logs
from a sample including a hundred projects prior to running the script
from Figure~\ref{fig:mvn-execution}.
The tasks we found were ignored from execution (lines 1-4).
Furthermore, we configured
our workstation to only run essential services as to avoid noise from unrelated OS events.
The machine was dedicated to our experiments and we
accessed it via SSH. In addition, we configured the \CodeIn{isolcpus}
option from the Linux Kernel \cite{linux-kernel} to isolate six
virtual CPUs to run our experiments, leaving the remaining CPUs to run
OS processes~\cite{isolcpus-use}.  The rationale for this decision is
to prevent context-switching between user processes (running the
experiment) and OS-related processes.  Finally, to make sure our
measurements were fair, we compared timings corresponding to the
sequential execution of tests using Maven with that obtained with
JUnit's default \CodeIn{JUnitCore} runner, invoked from the command
line.  Results were very close.
The main loop (lines 6-15) of the script in
Figure~\ref{fig:mvn-execution} iterates over the list of subjects and
invokes Maven multiple times\Comment{ to isolate cost of running
tests} (lines 8-11).  It first makes all dependencies available locally
(line 8), compiles the source and test files (line 9), and then runs
the tests in offline mode as to skip the package update task, enabled
by default (line 11). After that, we used a regular expression on
the output log to find elapsed times (line 12-14).

\input{codes/evaluation}

\begin{figure}[ht]%
  \vspace{-3mm}
  \centering
  \subfigure[\label{fig:rq1-barplot}]{
  \includegraphics[height=4cm,width=3cm]{results/barplot-timecost.pdf}
  }
  \qquad
  \subfigure[\label{fig:rq1-boxplot}]{
    \includegraphics[height=4cm,width=4.2cm]{results/boxplot-timecost.pdf}
  }%
  \vspace{-2mm}
  \caption{(a)\Comment{ Testing time grouped by time cost ($t$): short run
    ($t<1m$), medium run ($1m \le t < 5m$), and long run ($5m \le{}
    t$);}~Number of projects in each cost group and
    (b)~Distribution of running times per cost group.}
\vspace{-5mm}
\end{figure}

\Comment{We followed a similar methodology to group projects by time as
Gligoric~\etal{}~\cite{gligoric-etal-issta2015} in their work on
regression test selection.}
\Comment{ and added the \medg{} group due to the variability of the
time cost from subjects out of the \shortg{} group}

We ran the test suite for each subject three times, reporting averaged
execution times in three ranges: tests that run within a minute
(\shortg{}), tests that run in one to five minutes (\medg{}), and
tests that run in five or more minutes (\longg{}). Figure~\ref{fig:rq1-barplot} shows the number of projects in
each group.  As expected, \longg{} and \medg{} projects do not occur
as frequently as \shortg{} projects.  However, they do occur in
relatively high numbers.
Figure~\ref{fig:rq1-boxplot} shows the distribution of execution time of
test suites in each of these groups.
Note that the y-ranges are different.
The distribution associated with the \shortg{} group is the most
unbalanced (right skewed)\Comment{ with outliers closed to the \medg{}
group}.
The test suites in this group ran in 15 or less seconds for
over 75\% of the cases.\Comment{  Such scenarios constitute the majority of the
cases we analyzed.} Considering the groups \medg{} and \longg{},
however, we found many costly executions.  Nearly 75\% of the projects
from the \medg{} group take 3.5 or more minutes to run and nearly 75\% of
the projects from the \longg{} group take $\sim$20 minutes to run.  We
found cases in the \longg{} group were execution takes more than 50 minutes
to complete, as can be observed from the outliers in the boxplot.

It is important to note that we under-estimated running times as we missed test modules not enabled for
execution in the root \emph{pom.xml}.\Comment{Some projects may omit long-running tests on their default
execution.} For instance, the project \CodeIn{apache.maven-surefire}
runs all unit tests in a few seconds.  According to our criteria, this
project is classified as \shortg{} but a closer look reveals
that only smoke tests are executed in this project by default.
In this project, integration and system tests, which take longer to run, are only accessible via
custom parameters, which we do not handle in our experimental setup.
We enabled such parameters for this specific project and observed that
testing time goes to nearly 30 minutes.  For simplicity, we considered
only the tests executed by default.
From the \numSubjs{} testable projects, \numSubjsPass{} successfully
executed all tests and \numSubjsFail{} reported some test failure.
From these \numSubjsFail{} subjects, only 11 subjects
have more than 5\% of failing tests (7.3\% on average).
\begin{mdframed}
\noindent\textit{Answering \numRQFeasibilityOne{}:}~\emph{We conclude that
    time-consuming test suites are relatively frequent in
    open-source projects.  We found that \percentMedLongRunning{} of
    the \numSubjs{} projects we analyzed (\ie{}, nearly 1 in every 4
    projects) take at least 1 minute to run and
    \percentLongRunning{} of them take at least 5 minutes to run.\Comment{
      (\ie, \numMedLong{} projects from \medg{} and \longg{}).}}
\end{mdframed}

\begin{itemize}
  \item \numRQFeasibilityTwo. \textbf{\RQFeasibilityTwo}
\end{itemize}

\begin{figure}[t!]
  \centering
  \includegraphics[width=.48\textwidth]{results/testcost-distribution.pdf}
  \vspace{-2mm}
  \caption{\label{fig:time-distributions}Distribution of test case time per project.}%
  \vspace{-5mm}
\end{figure}

Section~\ref{sec:rqA} showed that medium and long-running projects are
not uncommon, accounting to nearly \percentMedLongRunning{} of the
\numSubjs{} projects we analyzed.  Research question \numRQFeasibilityTwo{}
measures the distribution of test costs in test suites.\Comment{ as to estimate
potential of obtaining speedups with parallelization.}  In
the limit, if cost is dominated by a single test from a large test
suite, it is unlikely that parallelization will be beneficial as a
test method is the smallest working unit in test frameworks.
Figure~\ref{fig:time-distributions} shows the time distribution of
individual test cases per project.
We observed that the average median times (see dashed horizontal red
lines) were small, namely 0.08s
for \medg{} projects and 0.16s for \longg{} projects, and the standard deviations associated with each distribution were
relatively low.\Comment{ Figure~\ref{fig:sd} shows the number of
  projects within specific ranges of $\sigma$ values.} High values of
$\sigma$ are indicative of CPU monopolization. We found only a small number
of those. The highest value of $\sigma$ occurred in
\CodeIn{uber\_chaperone}, a project from the \longg{} group.
This project contains only 26 tests, 17 of which take less than 0.5s
to run, one of which takes nearly 3s to run, two of which take nearly
11s to run, four of which takes on average 3m to run, and two of which
take $\sim$8m to run.
For this project, 98.4\% of the execution cost is dominated by 20\% of
the tests; without these two costly tests this project would have been
classified as short-running.
We did not find other projects with such extreme time monopolization
profile.
Project \CodeIn{facebookarchive\_linkbench} is also classified as
long-running and has the second highest value of $\sigma$.
For this project, however, cost is distributed more smoothly across 98
tests, of which 8 (8.1\%) take more than 1s to run with the rest of
the tests running faster.

\begin{figure}[t!]%
  \centering
  \subfigure[\label{fig:size-testsuites}]{
  \includegraphics[height=4.25cm, width=2.1cm]{results/boxplots-testcases.pdf}
  }
  \qquad
  \subfigure[\label{fig:scattercost}]{
    \includegraphics[height=4.25cm]{results/scatter-testcost.pdf}
  }%
  \vspace{-2mm}
  \caption{\label{fig:time-versus-size}(a) Size of test suites; (b)
    Size versus running time of test suites.}%
  \vspace{-5mm}
\end{figure}

Figure~\ref{fig:size-testsuites} shows the difference in the
distribution of test suite sizes across groups.  This figure indicates
that long projects have a higher median and much higher average number of test cases.
Furthermore, we noted a strong positive correlation between running
time and number of test on projects in the \longg{} group.
Considering the \medg{} group, the correlation between these two
variables was weak.
Figure~\ref{fig:scattercost} illustrates the regression lines between
these the variables test suite cost and number of test cases.
To sum, we observed that for projects with long-running test suites
running time is typically
justified by the number of test cases as opposed to the cost of individual test cases.

\begin{mdframed}
  \noindent\textit{Answering \numRQFeasibilityTwo{}:}~\emph{Overall, results indicate that
  projects with a very small number of tests monopolizing end-to-end
  execution time were rare. Time most often is distributed evenly
  across test cases.}
\end{mdframed}

\subsection{Adoption}
\label{sec:rqC}
\label{sec:rqE}

\begin{itemize}
    \item \numRQAdoptionOne. \textbf{\RQAdoptionOne{}}
\end{itemize}

To answer \numRQAdoptionOne{} we used projects from the \medg{} and
\longg{} groups where parallelization can be more helpful. We used a
dynamic and a static approach to find manifestations of
parallelism. We discuss results obtained with these complementary
approaches in the following.

\subsubsection{Dynamic checking}
\label{sec:rqC-1}

To find dynamic evidence of parallelism, we ran the test suites from
our set of \numMedLong{} projects to output all key-value pairs of
Maven parameters.  To that end, we used the option~\CodeIn{-X} to
produce debug output and the option~\CodeIn{-DskipTests} to skip
execution of tests.  We skipped execution of tests as we observed from
sampling that only bootstrapping the Maven process suffices to infer
which parallel configuration modes it uses to run the
tests.  It is also important to point that we used the default
configurations specified in the project.  We inferred parallel
configurations by searching for certain configuration parameters in
log files. According to Maven's
documentation~\cite{maven-surefire-plugin}, a parallel configuration
depends either on (1) the parameter \CodeIn{parallel} to define the
parallelism mode within a JVM followed by the parameter
\CodeIn{threadCount} or (2) the parameter
\CodeIn{forkCount}\footnote{This parameter is named \CodeIn{forkMode}
  in old versions of Maven Surefire.} to define the number of forked
JVMs.  As such, we captured, for each project, all related key-value
pairs of Maven parameters and mapped those pairs to one of the
possible parallelization modes.  For instance, if a given project
contains a module with the parameter
\CodeIn{<forkCount>1C</forkCount>}, the possible classifications are
\ForkSeq{} or \ForkParMeth{}, depending on the presence and the value
of the parameter \CodeIn{parallel}.  If the parameter
\CodeIn{parallel} is set to \CodeIn{methods} the detected mode will be
\ForkParMeth{}.  Large projects may contain several test suites
distributed on different Maven modules potentially using different
configurations.  For those cases, we collected the Maven output from
each module discarding duplicates as to avoid inflating counts for
configuration modes that appear in several modules of the same
project. For instance, if a project contains two modules using the
same configuration, we counted only one occurrence.
Considering
our set of \numMedLong{} projects, we found that only
\numProjectsPar{} of those projects had parallelism enabled
by default, with only configurations \ParClassSeqMeth{},
\ParClassParMeth{}, and \ForkSeq{} being used. Configurations
\ParClassParMeth{} and \ForkSeq{} were the most popular among these
cases. Note that these results under-approximate real usage of
parallelism as we used default parameters in our scripts to spawn the
Maven process.  That decision could prevent execution of particular
test modules. Table~\ref{tab:freqmodes-dynamic} shows the
\numProjectsPar{} projects we identified where parallelism is enabled by default in Maven.

Column ``\emph{Subject}'' indicates the name of the project, column
\begin{wraptable}{r}{5cm}
\vspace{-4mm}
\caption{Subjects with parallel test execution enabled by default.}
\label{tab:freqmodes-dynamic}
\footnotesize
\centering
\setlength{\tabcolsep}{2.5pt}
\begin{tabular}{llcr}
\toprule
\multirow{2}{*}{\emph{Group}} & \multirow{2}{*}{\emph{Subject}} & \emph{\# of} & \multirow{2}{*}{\emph{Mode}}\\%
   & & \emph{modules} &\\%
\midrule%
Medium & Californium & 2/20 & C2\\%
Medium & Chaos & 1/1 & C2\\%
Long & \Comment{apache} Flink & 66/74 & FC0\\%
Long & \Comment{apache logging-}Log4J2 & 25/28 & FC0\\%
Long & \Comment{javaslang }Javaslang & 3/3 & C3\\%
Medium & Jcabi \Comment{jcabi-github} & 1/1 & C3\\%
Long & \Comment{hazelcast hazelcast-}Jet & 6/7 & FC0\\%
Long & \Comment{apache} Mahout & 8/9 & FC0\\%
Long & \Comment{jankotek} MapDB & 1/1 & C3\\%
Medium & \Comment{apache} OpenNLP & 4/4 & FC0\\%
Medium & \Comment{yegor256} Rultor & 1/1 & C3\\%
Medium & \Comment{yegor256} Takes & 1/1 & C3\\%
Long & \Comment{vavr-io} Vavr & 3/3 & C3\\%
\bottomrule%
\end{tabular}
\end{wraptable}
``\emph{\# of modules}'' indicates the fraction of modules containing
tests that use the configuration of parallelism mentioned in column
``\emph{Mode}''.
We note that, considering these projects, the modules that do not use
the configuration cited use the sequential configuration \Seq{}.
For example, three modules (=28-25) from Log4J2 use sequential
configuration. It came as a surprise the observation that
no project used distinct configurations in their modules.

\subsubsection{Static checking}
\label{sec:rqC-2}
Given that the dynamic approach cannot detect parallelism manifested
through the default
configuration of projects, we also searched for indications of parallelism in build
files\Comment{ in the same sample set of \numMedLong{} projects}.  We
parsed all \emph{pom.xml} files under the project's directory and used
the same approach as in our previous analysis to classify
configurations.  We noticed initially that our approach was unable to
infer the configuration mode for cases where the decision depends on
the input (\eg,
\CodeIn{<parallel>\$\{parallel.type\}</parallel>}). For these
projects, the tester needs to provide additional parameters in the
command line to enable parallelization (\eg, \CodeIn{mvn test
  -Dparallel.type=classesAndMethods}). To handle those cases, we
considered all possible values for the parameter (in this case,
\CodeIn{\$\{parallel.type\}}).  It is also important to note that this
approach is not immune to false negatives, which can occur when
\emph{pom.xml} files are encapsulated in jar files or files downloaded from
the network.  Consequently, this approach complements the
the dynamic approach. Overall, we found \numProjectsParStatic{}
projects manifesting parallelism with this approach.
Compared to the set of projects listed in
Table~\ref{tab:freqmodes-dynamic}, we found four new projects, namely:
\CodeIn{Google Cloud\Comment{ Platform} DataflowJavaSDK} (using
configuration \ParClassParMeth), \CodeIn{Mapstruct} (using configuration
\ForkSeq{}), \CodeIn{T-SNE-Java} (using configuration \ForkSeq), and
\CodeIn{Urbanairship Datacube} (using configuration \ParClassParMeth).
Curiously, we also found that project \CodeIn{Jcabi}, \CodeIn{Rultor},
and \CodeIn{Takes} were not detected using this methodology.
That happened because these projects loaded a \emph{pom.xml} file from
a jar file that we missed.
Considering the static and dynamic methods together, we found a total
of \numProjectsParTotal{} distinct projects using parallelism,
corresponding to the union of the two subject sets.

\begin{mdframed}
  \noindent\textit{Answering \numRQAdoptionOne{}:}~\emph{Results indicate that test
  suite parallelization is underused.  Overall, only
  \percentParallel{} of costly projects (\numProjectsParTotal{} out of \numMedLong)
  use parallelism.}
\end{mdframed}

\begin{itemize}
  \item \numRQAdoptionTwo{}. \textbf{\RQAdoptionTwo{}}
\end{itemize}

To answer this research question we surveyed developers involved in a
selection of projects from our benchmark with time-consuming test
suites.  The goal of the survey is to better comprehend developer's
attitude towards the use of parallelism as a mechanism to speedup
regression testing.  We surveyed developers from a total of
\emailsProjects{} projects.  From the initial list of \numMedLong{}
project, we discarded 11 projects that we knew a priori used
parallelization, and \discartedProjects{} projects that we could not find
developer's emails from commit logs.  From this list of projects, we
mined potential participants for our study.  More precisely, we
searched for developer's name and email from the last 20 commits to
the corresponding project repository.  Using this approach, we
identified a total of \emailsSent{} eligible participants.  Finally,
we sent plain-text e-mails, containing the survey, to those developers.  In
total, \emailsAnswered{} developers replied but we discarded
\emailsFalseAnswers{} replies with subjective answers.  Considering
projects covered by the answers, a total of \emailsProjectsAnswered{}
projects (\percEmailsProjectsAnswered{} of the total) were represented
in those replies. Note that multiple developers on each project
received emails. In one specific case, one developer worked in
  multiple projects, and we consider it as a different answer. We sent the following set of questions to
developers:

\begin{enumerate}
\item How long does it take for tests to run in your environment? Can
  you briefly define your setup?
\item Do you confirm that your regression test suite does *not* run in parallel?
\item\label{questionThree} Select a reason for not using parallelization:
  \begin{enumerate}[label=\alph*)]
  \item I did not know it was possible
  \item I was concerned with concurrency issues
  \item I use a continuous integration server
  \item Some other reason. Please elaborate.
  \end{enumerate}
\end{enumerate}

Considering question 1, we confirmed that execution time was
compatible with the results we reported in Section~\ref{sec:rqA}.
Furthermore, \emailsCI{} of the participants indicated the use of
Continuous Integration (CI) to run tests, with \emailsDistributed{} of
these participants reporting that test suites are modularized and
those modules are tested independently in CI servers through different
parameters.  Those participants explained that such practice helps to
reduce time to observe test failures, which is the goal of speeding up
regression testing.  A total of \emailsLocal{} participants answered
that they do run tests in their local machines.  Note, however, that
CI does not preclude low-level parallelization.  For example,
installations of open-source CI tools (\eg{}, Jenkins~\cite{jenkins})
in dedicated servers would benefit from running tests faster through
low-level test suite parallelization.

Considering question 2, the answers we collected indicated, to our
surprise, that six of the \emailsProjectsAnswered{} projects execute
tests in parallel.  This mismatch is justified by cases where neither
of our checks (static or dynamic) could detect presence of
parallelism.  A closer look at these projects revealed that one of
them contained a \emph{pom.xml} file encapsulated in a jar file
(similar case as reported in Section~\ref{sec:rqC-2}), in one of the
projects the participant considered that distributed CI was a form of
parallelism, and in four projects the team preferred to implement
parallelization instead of using existing features from the testing
framework and the build system~---~in two projects the team
implemented concurrency control with custom JUnit test runners and in
two other projects the team implemented concurrency within test
methods.  Note that, considering these four extra cases (ignored two
distributed CI cases), the usage of parallelization increases from
\percentParallel{} to \percentParallelUpdated{}.  We do not consider
this change significant enough to modify our conclusion about
practical adoption of parallelization (\numRQAdoptionOne{}).

Considering question 3, the distribution of answers was as follows.  A
total of \emailsA{} of the \emailsProjectsAnswered{} developers who
answered the survey did not know that parallelism was available in
\begin{wrapfigure}{r}{0.13\textwidth}
%\begin{figure}[b]
  \centering
  \includegraphics[width=2cm]{results/survey.pdf}
  \caption{\label{fig:rq5-answers}Summary of developer's answers to
    survey question~\ref{questionThree}.}
\end{wrapfigure}
Maven (option ``a''), \emailsB{} of developers mentioned that they did
not use parallelism concerned with possible concurrency issues (option
``b''), \emailsD{} of developers mentioned that continuous integration
suffices to provide timely feedback while running only smoke
tests (\ie{}, short-running tests) locally (option ``c'')\Comment{here
  I want to say that they use it for something like "non-blocking
  testing" while developing in a local machine}, and \emailsD{} of
developers who provided an alternative answer (option ``d'') mentioned
that using parallelism was not worth the effort of preparing the test
suites to take advantage of available processing power.  A total of
\emailsNA{} of participants did not answer the last question of the
survey.  The pie chart in Figure~\ref{fig:rq5-answers} 
summarizes the distribution of answers.

\begin{mdframed}
\noindent\textit{Answering \numRQAdoptionTwo{}:}~\emph{Results suggest that dealing
  with concurrency issues (\ie{}, the extra work to organize test
  suite to safely explore concurrency) was the principal reason
  for developers not investing in parallelism.  Other reasons
  included availability of continuous integration services and
  unfamiliarity with the technology.}
\end{mdframed}

\subsection{Speedups}
\label{sec:rqD}

\begin{itemize}
    \item \numRQSpeedupOne{}. \textbf{\RQSpeedupOne}
\end{itemize}

To answer \numRQSpeedupOne{}, we considered the \numProjectsPar{}
subjects from our benchmark that use parallelization \emph{by default}
(see Table~\ref{tab:freqmodes-dynamic}).  We compared running times
of test suites with enabled parallelization, as configured by project
developers, and without parallelization. It is important to note that
there are no observed failures in either execution.
Table~\ref{tab:speedup} summarizes results.
Lines are sorted by project names.
Columns ``\emph{Group}'' and ``\emph{Subject}'' indicate, respectively,
the cost group and the name of the project.
Column ``$T_s$'' shows sequential execution time and column ``$T_p$''
shows parallel execution time.
Column ``$T_s/T_p$'' shows speedup or slowdown.
As usual, a ratio above 1x indicates speedup and a ratio below 1x
indicates slowdown.

Results show that, on average, parallel execution was
\avgSpeedup{} times faster compared to sequential execution.
Three cases worth special attention: \CodeIn{Log4J2}, \CodeIn{Chaos},
and \CodeIn{Takes}.
We note that parallel execution in \CodeIn{Log4J2} was
ineffective.  We found that Maven invokes several test modules in this
project but the test modules that dominate execution time run
sequentially by default. This was also the case for the highlighted
project \CodeIn{Californium}.
No significant speedup was observed in \CodeIn{Chaos}, a project with
only three test classes, of which one monopolizes the bulk of test
execution time.
This project uses configuration \ParClassSeqMeth{}, which runs test
classes in parallel but runs test methods, declared in each class,
sequentially.
Consequently, speedup cannot be obtained as the cost of the single
expensive test class cannot be broken down with the selected
configuration.
Finally, the speedup observed in project \CodeIn{Takes} was
the highest amongst all projects. This subject uses configuration
\ParClassParMeth{} and contains 419 test methods distributed nearly
equally among 148 test classes with a small number of test methods.
Furthermore, several methods in those classes are time-consuming.
As result, the CPUs available for testing are kept occupied for the
most part during test execution.

\begin{table}[t!]
\centering
\caption{
\label{tab:speedup}
Speedup (or slowdown) of parallel execution ($T_p$)
over sequential execution ($T_s$).  Default parallel configuration of
Maven is used.  Highest slowdown/speedup appears in gray color.}
\resizebox{.41\textwidth}{!}{%
%  \scriptsize
\begin{tabular}{llrrr}
\toprule
\emph{Group} & \emph{Subject} & \multicolumn{1}{c}{$T_s$} & \multicolumn{1}{c}{$T_p$} & $T_s/T_p$ \\%
\midrule%
Medium & Californium & 1.45m & 1.40m & \cellcolor{lightgray}1.04x\\%
Medium & \Comment{BounceStorage} Chaos\Comment{ HTTP Proxy} & 1.51m & 1.47m & \cellcolor{lightgray}1.03x\\%
Medium &\Comment{ Apache }Flink& 11.79m & 2.57m & 4.59x\\%
Long &\Comment{ Apache }Log4J2& 8.24m & 8.21m & \cellcolor{lightgray}1.00x\\%
Medium &Javaslang& 2.18m & 1.82m & 1.20x\\%
Medium &Jcabi\Comment{ GitHub} & 2.76m & 0.30m & 9.20x\\%
Long &\Comment{ Hazelcast }Jet& 8.26m & 3.67m & 2.25x\\%
Long & \Comment{apache} Mahout & 27.38m & 18.15m & 1.51x\\%
Long &\Comment{ Jankotek }MapDB& 10.06m & 8.58m & 1.17x\\%
Medium & \Comment{apache} OpenNLP & 1.30m & 0.55m & 2.36x\\%
Medium & \Comment{yegor256} Rultor & 2.30m & 0.27m & 8.52x\\%
Medium & \Comment{yegor256} Takes & 2.00m & 0.19m & \cellcolor{lightgray}10.53x\\%
Long & \Comment{vavr-io} Vavr & 3.26m & 2.25m & 1.45x\\%
\midrule
Average &  &  &  & \avgSpeedup{}x\\
\bottomrule%
\end{tabular}
}
\end{table}

\begin{mdframed}
 \noindent\textit{Answering \numRQSpeedupOne{}:}~\emph{Considering the
  machine setup we used, the average speedup observed with default
  configurations of parallelization was \avgSpeedup{}x.}
\end{mdframed}

\begin{itemize}
    \item \numRQSpeedupTwo{}. \textbf{\RQSpeedupTwo}
\end{itemize}

\newcommand{\subjectScalability}{MapDB}

This experiment evaluates the impact of making a growing number of
CPUs available to the build system for testing.  For this reason, we
used a different machine, with more cores, compared to the one described in
Section~\ref{sec:setup}.  We used a Xeon E5-2660v2 (2.20GHz) Intel
processor machine with 80 virtual CPUs (40 cores with two native
threads each) and 256GB of memory, running Ubuntu 14.04 LTS Trusty
Tahr (64-bit version). This experiment
spawns a growing number of JVMs in different CPUs, using parallel
configuration \emph{\ForkSeq{}}. We selected
subject \subjectScalability{} in this experiment as it represents the
\begin{wrapfigure}{r}{0.23\textwidth}
  \includegraphics[width=0.23\textwidth]{R/scalability/scalability.pdf}
  \caption{\label{fig:scalability}Scalability.}
  \vspace{-4mm}
\end{wrapfigure}
case of a long-running test suite (see Table~\ref{tab:speedup}) with
test cases distributed across many test classes~--~194.  
Recall that a test class is the smallest unit that can be used to spawn a test job
on a JVM and that we have no control over which test classes will be
assigned to which JVM that the build system forks.
Figure~\ref{fig:scalability} shows the reduction in running times as
more CPUs contribute to the execution.
We ran this experiment for a growing number of cores 1, 3, ..., 39. 
The plot omits results beyond 17 cores as the tendency for higher
values is clear.
We noticed that improvements are marginal after three cores, which is
the basic setup we used in other experiments.
This saturation is justified by the presence of a single test class,
\CodeIn{org.mapdb.WALTruncat}, containing 15 test cases that take over
two minutes to run.

\begin{mdframed}
\noindent\textit{Answering \numRQSpeedupTwo{}:}~\emph{Results suggest that
  execution FC0 scales with additional cores but there is a bound
  on the speedup that one can get related to how well the test suite is
  balanced across test classes.}
\end{mdframed}

\subsection{Tradeoffs}
\label{sec:rq6-tradeoffs}

This dimension assesses the impact of using distinct parallel
configurations on test flakiness and speedup.  Increased parallelism
can increase resource contention leading to concurrency issues such as
data races across dependent
tests~\cite{luo-etal-fse2014,bell-etal-esecfse2015}.  Flakiness and
speedup are contradictory forces that could influence the decision of
practitioners about which parallel configuration should be used for
testing.  Note that Section~\ref{sec:rqD} evaluated speedup in
isolation.

\begin{itemize}
  \item \numRQIssuesOne{}. \textbf{\RQIssuesOne{}}
\end{itemize}

\input{flakiness-speedup-table}

To answer this research question, we selected 15 different subjects,
ran their test suites against all configurations described in
Section~\ref{sec:modes}, and compared their running times and rate of
test flakiness.  We used the sequential execution configuration,
\emph{\Seq{}}, as the comparison baseline in this experiment.  To
select subjects, we sorted projects whose test suites run in 1m or
more by decreasing order of execution time and selected the first
fifteen projects that use JUnit 4.7 or later.  The rationale for this
criteria is to ensure compatibility with parallel configuration since
older versions of JUnit does not support parallel testing.  We ran
each project on each configuration for \SubjectsReruns{} times.
Overall, we needed to reran test suites 270 times, 18 times (3x6
configurations) on each project.  Given the low standard deviations
observed in our measurements\Comment{ and the aggregated time cost of 270 test
executions}, we considered \SubjectsReruns{} reruns reasonable for this
experiment.

It is worth mentioning that we used custom JUnit runners as opposed to
Maven to run the test suites with different parallel configurations
(see Section~\ref{sec:modes}).  After carefully checking library
versions for compatibility issues and comparing results with JUnit's
we observed that several of Maven's executions exposed problems.  For
example, Maven incorrectly counts the number of test cases executed
for some of the cases where test flakiness are observed. These issues are
categorized and documented on our website~\cite{ourwebpage} and can be
reproduced with our scripts.  To address those issues we implemented
custom test runners for configurations \emph{\SeqClassParMeth{}},
\emph{\ParClassSeqMeth{}}, and \emph{\ParClassParMeth{}} and,  for configurations
\emph{\ForkSeq{}} and \emph{\ForkParMeth{}}, we
implemented a bash script that coordinates the creation of JVMs and
invokes corresponding custom runners.  As to faithfully reflect
Maven's behavior in our scripts, we carefully analyzed the source
code~\cite{maven-surefire-source} of the Maven Surefire plugin. We
implemented test runners using the \CodeIn{ParallelComputer} class from
JUnit~\cite{junit-parallel}.

We used Maven log files to identify test classes to run and used the
Maven dependency plugin~\cite{maven-dep} to build the project's
classpath (with the command \CodeIn{mvn dependency:build-classpath}).
Once we find the tests suite to run and the corresponding classpath,
we invoke the test runners mentioned above on them.  We configured
this experiment to run at most three JVMs in parallel.  Recall that in
our setup (see Section~\ref{sec:setup}), we limited our kernel to use
only three cores and reserved one core for OS-related processes.  To
ensure that our experiments terminate (recall that deadlock or
livelock could occur) we used the \CodeIn{timeout}
command~\cite{timeout-cmd} configured to dispatch a \emph{kill} signal
if test execution exceeds a given time limit. Finally, we save each
execution log and stack traces generated from JUnit to collect the
execution time, the number of failing tests, and to diagnose outliers
in our results.

\sloppy Table~\ref{tab:rq6-table} summarizes results ordered by
subject's name.  Values are averaged across multiple executions.  We
did not report standard deviations as they are very small in all
cases.  As to identify the potential causes of flakiness, we inspected
the exceptions reported in execution logs. We found that, in most
of the cases, flakiness was caused by race conditions: approximately
97.5\% of the failures were caused by a \CodeInF{null} dereference and 1.6\% were
caused by concurrent access on unsynchronized data structures.\Comment{
  More precisely, the manifested exceptions were
  \CodeInF{ConcurrentModificationException} (0.8\%),
  \CodeInF{NoSuchElementException} (0.7\%), and
  \CodeInF{ArrayIndexOutOfBoundsException} (0.7\%).}  Cases of likely
broken test dependencies were not as prevalent as race conditions
(0.8\% of the total): \CodeInF{EOFException} (0.2\%),
\CodeInF{FileSystemAlreadyExistsException} (0.2\%), and
\CodeInF{BufferOverflowException} (0.4\%). Results suggest that
anticipating race conditions to schedule test executions would have
higher impact compared to breaking test dependencies using a tool such
as ElectricTest~\cite{bell-etal-esecfse2015}.

The projects with flakiness
in all configurations were \CodeIn{AWS SDK}, \CodeIn{GoogleCloud}, and
\CodeIn{Moquette}.  It is worth highlighting the unfortunate case of
\CodeIn{Moquette}, which manifested more than 20\% flaky tests in
every configuration.  Considering time, it is noticeable from the
averages, perhaps as expected, an increasing speedup from
configuration \emph{\SeqClassParMeth} to \emph{\ParClassParMeth} and
from configuration \emph{\ForkSeq} to \emph{\ForkParMeth}.  It is also
worth mentioning that some combinations manifested slowdown instead of
speedup.  Recall that parallel execution introduces the overhead of
spawning and managing JVMs and threads.  Overall, results show that
0\% of flakiness have been reported in 30 of the 75 (=5x15)
pairs of project and configuration we analyzed (40\% of the total).
In for 4 of the 15 projects flakiness was not manifested in 
any combination pairs.  We noticed with some surprise that the average speedup of
configuration \emph{\SeqClassParMeth} was higher compared to
\emph{\ForkParMeth} indicating that it is not always the case that
using more CPUs pays off. Important to note that the cost of spawning
new JVMs can be significant in \emph{\ForkParMeth}.

\begin{mdframed}
\noindent\textit{Answering \numRQIssuesOne{}:~Overall results indicate that the
  test suites of 73.33\% of the projects we analyzed could be run in
  parallel without manifesting any flaky tests.  In some of these
  cases, speedups were significant, ranging from 1x to 28.8x.}
\end{mdframed}

% SECTION 5

\section{Discussion}

This paper reports our finding on a study to evaluate impact and usage
of test suite parallelization, enabled by modern build systems and
testing frameworks.  This study is important given the importance to
speedup testing.  Note that test suite parallelization is
complementary to alternative approach to speedup testing
(see~Section~\ref{sec:related}).  The observations we made in this
study trigger multiple actions:

\begin{itemize}
\item \emph{Incentivize forking.}~Forked JVMs manifest low rates of
  test flakiness.  For instance, in \emph{\ForkSeq{}}, only 4 of 10
  projects manifest flakiness and, excluding the extreme case of
  \CodeIn{Moquette}, projects manifest flaky tests in low rates 0.23\%
  to 1.70\%.  Developers of projects with long-running test suites
  should consider using that feature, which is available in modern
  build systems today (\eg{}, Maven).
\item \emph{Break test dependencies.}~Non-forked JVMs can achieve
  impressive speedups at the expense of sometimes impressive rates of
  flakiness.  Breaking test dependencies to avoid flakiness and take
  full advantage of those options is advised for developers with
  greater interest in efficiency.
\item \emph{Refactor tests for load balancing.}~Forked JVMs scales
  better with the number of cores when the test workload is balanced
  across testing classes.  To balance the workload, automated
  user-oblivious refactoring can help in scenarios where developers
  are not willing to change test code but have access to machines with
  a high number of cores.
\item \emph{Improve debugging for build systems.}~While preparing our
  experiments, we found scenarios\Comment{, related to test
    parallelization,} where Maven's executions did not reflect
  corresponding JUnit's executions. \Comment{(Docker reproduction scripts
  available.)} Those issues can hinder developers from using parallel
  testing. Better debugging infrastructure is important.
\end{itemize}

This study brings to light the benefits and burdens of test suite
parallelization to improve test efficiency. It provides
recommendations to practitioners and developers of new techniques and
tools aiming to speed up test execution with parallelization.

% SECTION 6
%%

\section{Threats to Validity}

The main threats to validity of this study are the following.

\textit{External Validity:} Generalization of our findings is limited
to our selection of subjects, testing framework, and build system.  To
mitigate that issue, we selected subjects according to an objective
criteria, described in Section~\ref{sec:subjects}.  It remains to
evaluate the extent to which our observations would change when using
different testing frameworks and build systems.
Also, some of the selected subjects contain failing tests. Test
failures may reduce the testing time due to early termination
or even inflate the time (\eg, test waiting indefinitely for
an unavailable resources).
To mitigate this threat, we eliminated subjects with flaky tests and
filtered projects with at least 90\% of the tests passing.
Only 17\% of our subjects have failing tests.
We carefully inspected our rawdata to identify and ignore these
failures with JUnit's \CodeIn{@Ignore} annotation.

\textit{Internal Validity:} Our results could be influenced by
unintentional mistakes made by humans who interpreted survey data and
implemented scripts and code to collect and analyze the data.
For example, we developed JUnit runners to reproduce Maven's parallel
configurations and implemented several scripts to automate our
experiments (\eg, run tests and detect parallelism enabled by default
in the subjects).
All those tasks could bias our results.
To mitigate those threats, the first two authors of this paper
validated/inspected each other to increase chances of capturing
unintentional mistakes.

\textit{Construct Validity:} We considered a number of metrics in this
study that could influence some of our interpretations.  For example,
we measured number of test cases per suite, distribution of test costs
in a suite, time to run a suite, etc.  In principle, these metrics may
not reflect the main problems associated with test
efficiency.

% SECTION 7
%%

\section{Related Work}
\label{sec:related}

Regression testing research has focused mostly on test suite
minimization, prioritization, reduction, and
selection~\cite{yoo-harman-stvr2012,soetens-etal-2016}.  Most of these
techniques are unsound (\ie{}, they do not guarantee that
fault-revealing tests will be considered for testing).  The test
selection technique
Ekstazi~\cite{gligoric-etal-issta2015,celik-etal-fse2017} is an
example of a sound regression testing technique. It conservatively
computes which tests have been impacted by file changes.  A test is
discarded for execution if it does not depend on any changed file
dynamically reachable from execution.\Comment{ Curiously Ekstazi's
  evaluation discovered subjects with parallelism enabled by
  default.}\Comment{ \c{C}elik~\etal{}~\cite{} recently extended
  Ekstazi to track files accessed outside JVM boundaries.} Important
to note that regression testing techniques, including test selection,
is complementary to test suite parallelization.

ElectricTest~\cite{bell-etal-esecfse2015} is a tool for efficiently
detecting data dependencies across test cases.  Dependency tracking is
important as to avoid test flakiness when parallelizing test
suites. ElectricTest observes reads and writes on global resources
made by tests to identify these dependencies at low cost. We remain to
investigate the impact of ElectricTest to reduce flakiness in
unrestricted test suite parallelization.

The use of the Single Instruction Multiple Data (SIMD) design has been
previously explored in research to accelerate test
execution~\cite{damorim-etal-issta2007,damorim-etal-tse2008,kim-etal-issre2012,nguyen-etal-icse2014,rajan-etal-ase2014,sen-etal-fse2015,yaneva-etal-issta2017}. The
SIMD architecture, as implemented in modern GPUs, for instance, allows
the execution of a given instruction simultaneously against multiple
data.  For that reason, in principle, one test could be ran
simultaneously against multiple inputs provided that multiple test
inputs exist associated to that one test.  Recent
work~\cite{rajan-etal-ase2014,yaneva-etal-issta2017} explored that
idea to speedup test execution of embedded software using graphic
cards. Although benchmarks indicate superior performance compared to
traditional multicore CPUs, the use of the technology in broader
settings is limited. For example, execution of more general programs
can violate the SIMD's lock-step assumption on the control-flow of
threads.  This violation would affect negatively performance.
Furthermore, handling complex data is challenging in
SIMD~\cite{damorim-etal-issta2007,damorim-etal-tse2008}.  The approach
is promising when multiple input vectors exist for each test and the
testing code heavily manipulates scalar data types.  The datasets used
in those papers satisfied those constraints.

Google~\cite{google-tap,google-ci} and
Microsoft~\cite{prasad-shulte-ieee-microsoft-ci} have been creating
distributed infrastructures to efficiently build massive amounts of
code and run massive amounts of tests.  Those scenarios bring
different and challenging problems such as deciding when to trigger
the build under multiple file
updates~\cite{memon-etal-icse17}. Although such distributed systems
are targeted to extremely large scale code and test bases, the same
ideas can be applied to handle the build process of large, albeit not
as large, projects.  For example,
Gambi~\etal{}~\cite{gambi-etal-issta2017} recently proposed CUT, a
tool to automatically parallelize JUnit tests on the cloud. The tool allows
the developer to control resource allocation and deal with the project
specific test dependencies.  Note that test suite parallelization is
complementary to these high-level parallelism schemes.

Continuous Integration (CI) services, such as Travis CI~\cite{travis},
are becoming widely used in the open-source
community~\cite{hilton-etal-ase2016,vasilescu-etal-fse2015}. Accelerating
time to run tests in CI is important as to reduce the period between
test report updates.  Module-level regression
testing~\cite{vasic-etal-fse2017}, for example, can be helpful in that
setting. It is important to note that test failures are more common in
CI compared to an overnight run or a local run, for instance.  This
can happen because of semantic merge conflicts~\cite{brun-etal-fse11},
for instance.  As such effect can impact developer's perception and
tolerance towards failures, we are curious to know if developers would
be willing to receive more frequent test reports at the expense of
potentially increasing failure rates due to flakiness caused by
parallelism.

% SECTION 8
%%

\section{Conclusions}

Testing is expensive.  Despite all advances in regression testing
research, dealing with high testing costs remains an important problem
in Software Engineering.  This paper reports our findings on the usage
and impact of test execution parallelization in open-source projects.
Multicore CPUs are widely available today.  Testing frameworks and
build systems that capitalize on these machines also became popular.
Despite some resistance observed from practitioners, our results
suggest that parallelization can be used in many cases without
sacrificing reliability. More research needs to be done to improve
automation (\eg{}, breaking test dependencies, refactoring test
suites, enforcing safe test schedules) as to safely optimize parallel
execution.  The artifacts we produced as result of this study are
available from the following web page:

\begin{center}
  \webpage{}
\end{center}

%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

% use section* for acknowledgment
%\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
%  \section*{Acknowledgments}
%\else
  % regular IEEE prefers the singular form
%  \section*{Acknowledgment}
%\fi

%The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\balance
\bibliographystyle{plain}
\bibliography{tmp}

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% that's all folks
\end{document}