### 2019-08-14 tse19/mitigating-flakiness-speedup-table.tex

%% Table header macros
% $\Uparrow_\text{speed}$

\begin{table*}[t]
\centering
\setlength{\tabcolsep}{4.5pt}
\caption{Speedup versus Flakiness (\subcolB). Configuration
  \emph{\Seq{}} denotes the comparison baseline.  Columns $T$
and $N$ indicate time and number of tests, respectively.  Other columns show
speedup and percentage of failing tests in different configurations, compared to
\emph{\Seq{}}.}
\begin{tabular}{l|rr|rr|rr|rr|rr|rr}
\toprule
\multirow{2}{*}{\emph{Subject}} & \multicolumn{2}{c|}{\emph{\Seq}} &
    \colheader{\SeqClassParMeth} & \colheader{\ParClassSeqMeth} &
    \colheader{\ParClassParMeth} & \colheader{\ForkSeq} &
    \colheader{\ForkParMeth} \\ %\cline{2-12}
    & $T$ & $\mathit{N}$ & \subcol{} & \subcol{} & \subcol{} & \subcol{}
    & \subcol{}\\%
\midrule%

Activiti & \entry{8.46m}{2093}  & \textcolor{blue}{ 2.49x } & \textcolor{blue}{ 0\%} & \textcolor{blue}{ 1.81x } & \textcolor{blue}{ 0\%} & \textcolor{blue}{ 1.72x } & \textcolor{blue}{ 0\%} & \textcolor{blue}{ 1.02x } & \textcolor{blue}{ 0\%} & \textcolor{blue}{ 2.87x } & \textcolor{blue}{ 0\%}\\%

Apache Avro & \entry{13.65m}{3765}  & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%}\\%

Apache Incubator Dubbo & \entry{14.33m}{2663}  & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%}\\%

Content ExtRactor and MINEr & \entry{8.10m}{132}  & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%} & \entry{-x}{-\%}\\%

DataStax Java Driver & \entry{5.23m}{1377}  & \entry{8.30x}{1.17\%} & \entry{9.34x}{0.75\%} & \entry{9.33x}{1.15\%} & \entry{1.03x}{0\%} & \entry{6.22x}{1.41\%}\\%

DROID & \entry{6.00m}{386}  & \entry{2.32x}{15.79\%} & \entry{7.90x}{2.02\%} & \entry{4.02x}{5.33\%} & \entry{9.67x}{2.3\%} & \entry{4.76x}{6\%}\\%

Eclipse Hawkbit & \entry{17.28m}{1111}  & \entry{2.63x}{29.12\%} & \entry{2.90x}{33.22\%} & \entry{2.52x}{26.52\%} & \entry{2.49x}{9.23\%} & \entry{3.00x}{33.91\%}\\%

Facebook Linkbench & 7.16m & 98 & 1.02x & 0\% & 1.70x & 1.27\% & 1.00x & 1.00\% & 1.70x & 0.09\% & 1.69x & 0.93\%\\%

Kafka Streams & \entry{11.38m}{79}  & \entry{1.13x}{11\%} & \entry{1.02x}{0\%} & \entry{1.12x}{11.49\%} & \entry{2.63x}{1.37\%} & \entry{1.02x}{0.2\%}\\%

Rapidoid & \entry{11.68m}{476}  & \entry{1.06x}{0\%} & \entry{1.24x}{0\%} & \entry{1.10x}{0\%} & \entry{18.54x}{10.15\%} & \entry{17.17x}{9.57\%}\\%

Search Guard & \entry{13.32m}{111}  & \entry{1.78x}{0\%} & \textcolor{blue}{ 1.62x } & \textcolor{blue}{ 0\%} & \textcolor{blue}{ 1.43x } & \textcolor{blue}{ 0\%} & \entry{2.04x}{0\%} & \entry{2.45x}{0\%}\\%

T SNE Java & \entry{6.38m}{20}  & \textcolor{blue}{ 2.02x } & \textcolor{blue}{ 0\%} & \entry{1.00x}{0\%} & \textcolor{blue}{ 1.22x } & \textcolor{blue}{ 0\%} & \entry{1.00x}{0\%} & \textcolor{blue}{ 1.21x } & \textcolor{blue}{ 0\%}\\%

Twilio Java & \entry{12.60m}{2529}  & \entry{1.47x}{1.37\%} & \entry{2.15x}{21\%} & \entry{3.08x}{2.18\%} & \entry{3.47x}{0\%} & \entry{2.60x}{5.54\%}\\%

Urban Airship Datacube & 11.84m & 45 & 2.31x & 20.91\% & 3.35x & 34.06\% & 1.92x & 21.46\% & 1.21x & 0\% & 2.26x & 21\%\\%

Vaadin Framework & \entry{8.97}{3683}  & \textcolor{blue}{ 2.19x } & \textcolor{blue}{ 0\%} & \textcolor{blue}{ 1.91x } & \textcolor{blue}{ 0\%} & \textcolor{blue}{ 2.01x } & \textcolor{blue}{ 0\%} & \entry{1.00x}{0\%} & \textcolor{blue}{ 2.53x } & \textcolor{blue}{ 0\%}\\%

\midrule
Average & 10.42m & 1,237.86 & -x & -\% &  -x & -\%  & -x  & -\% & -x & -\% & -x & -\%\\%

\bottomrule%
\end{tabular}
\label{tab:rq6-table}
\end{table*}