### 2019-09-10 tse19/mitigating-flakiness-speedup-table.tex
fix table

%% Table header macros
% $\Uparrow_\text{speed}$

\begin{table*}[t]
\centering
\setlength{\tabcolsep}{4.5pt}
\caption{Speedup versus Flakiness (\subcolB). Configuration
  \emph{\Seq{}} denotes the comparison baseline.  Columns $T$
and $N$ indicate time and number of tests, respectively.  Other columns show
speedup and percentage of failing tests in different configurations, compared to
\emph{\Seq{}}.}
\begin{tabular}{l|rr|rr|rr|rr|rr|rr}
\toprule
\multirow{2}{*}{\emph{Subject}} & \multicolumn{2}{c|}{\emph{\Seq}} &
    \colheader{\SeqClassParMeth} & \colheader{\ParClassSeqMeth} &
    \colheader{\ParClassParMeth} & \colheader{\ForkSeq} &
    \colheader{\ForkParMeth} \\ %\cline{2-12}
    & $T$ & $\mathit{N}$ & \subcol{} & \subcol{} & \subcol{} & \subcol{}
    & \subcol{}\\%
\midrule%

Activiti & \entry{8.46m}{2093}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%}\\%

Apache Avro & \entry{13.65m}{3765}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{0x}{0\%}\\%

Apache Incubator Dubbo & \entry{14.33m}{2663}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%}\\%

Content ExtRactor and MINEr & \entry{8.10m}{132}  & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{0x}{0\%}\\%

DataStax Java Driver & \entry{5.23m}{1377}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{-----}{-----} & \entry{0x}{0\%}\\%

DROID & \entry{6.00m}{386}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%}\\%

Eclipse Hawkbit & \entry{17.28m}{1111}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%}\\%

Facebook Linkbench & 7.16m & 98 & ----- & ----- & 0x & 0\% & 0x & 0\% & 0x & 0\% & 0x & 0\%\\%

Kafka Streams & \entry{11.38m}{79}  & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%}\\%

Rapidoid & \entry{11.68m}{476}  & \entry{-----}{-----} & \entry{-----}{-----} & \entry{-----}{-----} & \entry{0x}{0\%} & \entry{0x}{0\%}\\%

Search Guard & \entry{13.32m}{111}  & \entry{-----}{-----} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{-----}{-----}\\%

T SNE Java & \entry{6.38m}{20}  & \entry{1.26x}{0\%} & \entry{-----}{-----} & \entry{1.24x}{0\%} & \entry{-----}{-----} & \entry{1.23x}{0\%} \\%

Twilio Java & \entry{12.60m}{2529}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{0x}{0\%}\\%

Urban Airship Datacube & 11.84m & 45 & 0x & 0\% & 0x & 0\% & 0x & 0\% & ----- & ----- & 0x & 0\%\\%

Vaadin Framework & \entry{8.97}{3683}  & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{0x}{0\%} & \entry{-----}{-----} & \entry{0x}{0\%}\\%

\midrule
Average & 10.42m & 1,237.86 & ----- & ----- &  ----- & -----  & -----  & ----- & ----- & ----- & ----- & ----- \\%

\bottomrule%
\end{tabular}
\label{tab:rq6-table}
\end{table*}